---
title: "Ethnoracial Disparities in Disciplinary Action"
subtitle: "<b> Racial and Ethnic Disparities in Disciplinary Action in Schools, Depressive Symptoms, Substance Use and Academic Performance: Results from the ABCD Study </b><br> This document outlines the key stages in analyses, specifically it:<br> 
&nbsp;&nbsp;&nbsp;&nbsp;1. Performs data wrangling and runs creates the matched data<br>
&nbsp;&nbsp;&nbsp;&nbsp;2. Runs regression analyses<br>
&nbsp;&nbsp;&nbsp;&nbsp;3. Builds visualizations."

format:
  lumo-html: 
    primary-color: "#b5cef7"
    github-repo: "https://github.com/mayaogonah" 
    self-contained: true
    is-particlejs-enabled: true  # optional: display interactive particles in the doc header
    #bg-image:  # optional: add a background image
    code-fold: show

execute:
  eval: false
  echo: true
author: Siqi Wang (updated for ABCD 6.1 data release by Maya Ogonah)
date: today
---

------------------------------------------------------------------------

# Data Wrangling and Create the Matched Data

## Step 0: Install Required Packages (Run Once)

```{r install-packages, eval=FALSE}
# ============================================================
# RUN THIS CHUNK ONLY ONCE when setting up on a new computer.
# It installs all the R packages (add-on tools) this project needs.
# After packages are installed, you do NOT need to run this again.
# ============================================================

install.packages(
  c(
    'dplyr',      # Tools for filtering, selecting, and transforming data
    'purrr',      # Tools for applying functions repeatedly (used in merge)
    'tidyr',      # Tools for reshaping data (wide <-> long format)
    'ggplot2',    # Creating charts and plots
    'Hmisc',      # Descriptive statistics (describe() function)
    'gtsummary',  # Publication-ready summary tables
    'MatchIt',    # Propensity score matching
    'mice',       # Multiple imputation for missing data
    'VIM',        # Visualizing missing data patterns
    'table1',     # Table 1 demographic summary tables
    'flextable',  # Formatting tables for Word export
    'officer',    # Saving formatted tables to Word (.docx) files
    'hrbrthemes', # data visualisation themes
    'cowplot',    # combining plots
    'report',     # reporting ANOVA results
    'cobalt',     # assessing covariate balance
    'ggalluvial'  # adds geom_alluvium() and geom_stratum() to ggplot2 for Sankey/alluvial plots
  ),
  repos = 'https://cran.r-project.org/'
)
```

Install `renv` and initialise the project (run once). The `renv` package creates an isolated, project specific environment that records the packages and their versions used in the script.

```{r renv, eval=FALSE}
install.packages("renv")
renv::init()

# to update the project dependencies
renv::snapshot()
```

------------------------------------------------------------------------

## Step 1: Load Packages

Load all required packages into R's memory.

```{r load-packages, message=FALSE, warning=FALSE}
library(dplyr)      # Data manipulation
library(purrr)      # Functional programming (used in reduce() for merging)
library(tidyr)      # Data reshaping (pivot_wider, drop_na, etc.)
library(ggplot2)    # Plotting
library(Hmisc)      # Descriptive statistics
library(gtsummary)  # Summary tables
library(naniar)     # Missing data
library(VIM)        # Missing data patterns
library(MatchIt)    # Propensity score matching
library(mice)       # Multiple imputation
library(VIM)        # Missing data visualization
library(table1)     # Table 1 summaries
library(flextable)  # Formatted tables
library(officer)    # Word document export
library(stringr)    # ~ str_remove() function
library(hrbrthemes) # data visualisation themes
library(cowplot)    # combining plots
library(cobalt)     # assessing covariate balance
library(report)     # reporting ANOVA results 
library(ggalluvial) # adds geom_alluvium() and geom_stratum() to ggplot2 for Sankey/alluvial plots
```

------------------------------------------------------------------------

## Step 2: Create Data Visualisation Theme

These custom functions create a dataviz theme using the Klimt palette from the `MetBrewer` package.

```{r data-vis}
scale_color_brand <- function() {
  custom_palette <- MetBrewer::met.brewer(name = "Klimt", n=25, type="continuous") # <1>
  scale_color_manual(values = custom_palette)
}

scale_fill_brand <- function(picks = NULL, labels = NULL) {
  custom_palette <- MetBrewer::met.brewer(name = "Klimt", n=25, type="continuous")
  if (!is.null(picks)) {
    custom_palette <- custom_palette[picks]
  }
  if (is.null(labels)) {
    scale_fill_manual(values = custom_palette)
  } else {
    scale_fill_manual(values = custom_palette, labels = labels)
  }
}

palette <- MetBrewer::met.brewer(name = "Klimt", n=25, type="continuous")

theme_brand <- function() {
  theme_ipsum() + # <2>
    theme(
      plot.title = element_text(color = "black", size = 14, face = "bold"),
      plot.subtitle = element_text(margin = margin(b = 20)),
      axis.text.x = element_text(size = 7),
      axis.text.y = element_text(size = 7),
      axis.title.x = element_text(hjust = 0.5),  # center x-axis label
      axis.title.y = element_text(hjust = 0.5),  # center y-axis label
      plot.background = element_rect(fill = "gray97", color = NA),
      plot.caption = element_text(hjust = 0)
    )
}

theme_colors <- MetBrewer::met.brewer(name = "Klimt", n = 25, type = "continuous")
```

1.  The n function allows you to extend the number of values so the colour palette can still be used on graphs with numerous categorical values.
2.  You can change the font using the hrbrthemes package. I don't have the other fonts installed, and its a bit of a pain to install them, hence why I used theme_ipsum.

```{r}
scales::show_col(MetBrewer::met.brewer(name = "Klimt", n = 25, type = "continuous"))
```

------------------------------------------------------------------------

## Step 3: Define Helper Functions

This custom function pivots a LONG-format dataset to WIDE format.

ARGUMENTS:

1\. dataset - a data frame with columns: participant_id, session_id, \[column\]

2\. column - the name of the variable to pivot (as a string)

3\. targeted_outcome - optional; if provided, also creates an "ever\_\[column\]" column that is 1 if the participant EVER had that value

OUTPUT: a wide-format data frame with one row per participant

```{r define-functions}
pivot_wide <- function(dataset, column, targeted_outcome = NULL) {

  # Validate that the required columns exist
  if (!("eventname" %in% colnames(dataset))) {
    stop("The dataset must contain a column named 'eventname'.")
  }
  if (!("src_subject_id" %in% colnames(dataset))) {
    stop("The dataset must contain a column named 'src_subject_id'.")
  }
  if (!(column %in% colnames(dataset))) {
    stop(paste("The column", column, "is not found in the dataset."))
  }

  # If targeted_outcome is specified, create an "ever" indicator column.
  # Example: if column = "suspension" and targeted_outcome = 1,
  # then ever_suspension = 1 if the participant was EVER s%uspended.
  if (!is.null(targeted_outcome)) {
    ever_column_name <- paste0("ever_", column)

    dataset <- dataset %>%
      group_by(src_subject_id) %>%
      mutate(
        !!ever_column_name := ifelse(
          all(is.na(.data[[column]])), NA,
          as.integer(any(.data[[column]] == targeted_outcome, na.rm = TRUE))
        )
      ) %>%
      ungroup()
  }

  # Pivot from long to wide: each timepoint becomes a separate column
  # Column names will be like: [variable]_baseline_year_1_arm_1, etc.
  wide_dataset <- dataset %>%
    pivot_wider(
      id_cols     = src_subject_id,
      names_from  = eventname,
      values_from = c(column),
      names_glue  = "{.value}_{eventname}"
    )

  # If an "ever" column was created, merge it back onto the wide dataset
  if (!is.null(targeted_outcome)) {
    wide_dataset <- wide_dataset %>%
      left_join(
        dataset %>%
          select(src_subject_id, !!sym(ever_column_name)) %>%
          distinct(),
        by = "src_subject_id"
      )
  }

  return(wide_dataset)
}
```

------------------------------------------------------------------------

## Step 4: Load Raw Data

Load all required raw data from the ABCD Study dataset.

```{r load-data}

DEMO_COLS    <- c("participant_id", "session_id", "site",
                  "ab_g_stc__cohort_ethnrace__mhisp", "ab_g_stc__cohort_ethn",
                  "ab_g_dyn__visit_age", "ab_g_stc__cohort_sex",
                  "ab_g_dyn__cohort_income__hhold__6lvl", "ab_p_demo__marital__slf_001",
                  "ab_g_dyn__cohort_edu__cgs", "ab_p_demo__race_001___11")

KBI_COLS     <- c("participant_id", "session_id",
                  "mh_p_kbi__school_003", "mh_p_kbi__school_004")

SEDA_COLS    <- c("participant_id", "session_id",
                  "le_l_sedarec__distr__urban_indicator__2019",
                  "le_l_urban__addr1_urbanclassification")

CRPBI_COLS   <- c("participant_id", "session_id",
                  "fc_y_crpbi__cg2_000")

CBCL_COLS    <- c("participant_id", "session_id",
                  "mh_p_cbcl__synd__ext_tscore", "mh_p_cbcl__dsm__dep_tscore",
                  "mh_p_cbcl__synd__rule_tscore", "mh_p_cbcl__synd__rule_sum",
                  "mh_p_cbcl__synd__int_tscore", "mh_p_cbcl__rule__cond_002",
                  "mh_p_cbcl__aggr__opp_003")

TLFB_COLS    <- c("participant_id", "session_id",
                  "su_y_tlfb__nic_ud", "su_y_tlfb_ud")

SAG_COLS     <- c("participant_id", "session_id",
                  "fc_p_sag_005")

SUI_COLS     <- c("participant_id", "session_id")

abcd_all <- readxl::read_excel("data/7.0-data.xlsx")

  p_demo            <- abcd_all[, intersect(DEMO_COLS,  colnames(abcd_all))]
  mh_p_ksads_bg     <- abcd_all[, intersect(KBI_COLS,   colnames(abcd_all))]
  led_l_seda_demo_d <- abcd_all[, intersect(SEDA_COLS,  colnames(abcd_all))]
  ce_y_crpbi        <- abcd_all[, intersect(CRPBI_COLS, colnames(abcd_all))]
  cbcl              <- abcd_all[, intersect(CBCL_COLS,  colnames(abcd_all))]
  su_y_tlfb         <- abcd_all[, intersect(TLFB_COLS,  colnames(abcd_all))]
  ce_p_sag          <- abcd_all[, intersect(SAG_COLS,   colnames(abcd_all))]
  su_y_sui          <- abcd_all[, intersect(SUI_COLS,   colnames(abcd_all))]

```

------------------------------------------------------------------------

## Step 4b: Rename Variable names

The new ABCD release uses different variable names from 5.1. We rename them here once so that ALL code from Step 5 onward can continue using the original names unchanged.

```{r rename-to-internal-names}
# ============================================================
# RENAME NEW RELEASE VARIABLE NAMES             -> INTERNAL NAMES
#
# New release name                              -> Internal name used in code
# ---------------------------------------------------------------
# participant_id                                -> src_subject_id
# session_id                                    -> eventname
# ab_g_stc__cohort_ethnrace__mhisp              -> race_ethnicity
# ab_g_stc__cohort_ethn                         -> demo_ethn_v2
# ab_g_dyn__visit_age                           -> demo_brthdat_v2
# ab_g_stc__cohort_sex                          -> demo_sex_v2
# ab_g_dyn__cohort_income__hhold__6lvl          -> demo_prnt_income_v2
# ab_p_demo__marital__slf_001                   -> demo_prnt_marital_v2
# ab_g_dyn__cohort_edu__cgs                     -> demo_prnt_ed_v2_2yr_l
# ab_p_demo__race_001___11                      -> demo_race_a_p___11
# mh_p_kbi__school_003                          -> kbi_p_c_det_susp_l
# mh_p_kbi__school_004                          -> kbi_p_c_school_setting
# le_l_urban__addr1_urbanclassification         -> ledsch_seda_d_urban
# fc_y_crpbi__cg2_000                           -> crpbi_caregiver1_y
# mh_p_cbcl__synd__ext_tscore                   -> cbcl_scr_syn_external_t
# mh_p_cbcl__dsm__dep_tscore                    -> cbcl_scr_dsm5_depress_t
# mh_p_cbcl__synd__rule_tscore                  -> cbcl_scr_syn_rulebreak_t
# mh_p_cbcl__synd__rule_sum                     -> cbcl_scr_syn_rulebreak_sum
# mh_p_cbcl__synd__int_tscore                   -> cbcl_scr_syn_internal_t
# mh_p_cbcl__rule__cond_002                     -> cbcl_q28_p
# mh_p_cbcl__aggr__opp_003                      -> cbcl_q23_p
# su_y_tlfb__nic_ud                             -> su_tlfb_cal_scr_nic_days_yr
# su_y_tlfb_ud                                  -> tlfb_cal_scr_tot_su_ud
# fc_p_sag_005                                  -> sag_grade_type
#
# rename_if_present() only renames a column if it actually exists
# -- safe to run on both old and new release files.
# ============================================================

rename_if_present <- function(df, old, new) {
  if (old %in% colnames(df)) df <- rename(df, !!new := !!sym(old))
  df
}

apply_renames <- function(df) {
  df %>%
    rename_if_present("participant_id",                        "src_subject_id")              %>%
    rename_if_present("session_id",                            "eventname")                   %>%
    rename_if_present("ab_g_stc__cohort_ethnrace__mhisp",      "race_ethnicity")              %>%
    rename_if_present("ab_g_stc__cohort_ethn",                 "demo_ethn_v2")                %>%
    rename_if_present("ab_g_dyn__visit_age",                   "demo_brthdat_v2")             %>%
    rename_if_present("ab_g_stc__cohort_sex",                  "demo_sex_v2")                 %>%
    rename_if_present("ab_g_dyn__cohort_income__hhold__6lvl",  "demo_prnt_income_v2")         %>%
    rename_if_present("ab_p_demo__marital__slf_001",           "demo_prnt_marital_v2")        %>%
    rename_if_present("ab_g_dyn__cohort_edu__cgs",             "demo_prnt_ed_v2_2yr_l")       %>%
    rename_if_present("ab_p_demo__race_001___11",              "demo_race_a_p___11")          %>%
    rename_if_present("mh_p_kbi__school_003",                  "kbi_p_c_det_susp_l")          %>%
    rename_if_present("mh_p_kbi__school_004",                  "kbi_p_c_school_setting")      %>%
    rename_if_present("le_l_urban__addr1_urbanclassification", "ledsch_seda_d_urban")         %>%
    rename_if_present("fc_y_crpbi__cg2_000",                   "crpbi_caregiver1_y")          %>%
    rename_if_present("mh_p_cbcl__synd__ext_tscore",           "cbcl_scr_syn_external_t")     %>%
    rename_if_present("mh_p_cbcl__dsm__dep_tscore",            "cbcl_scr_dsm5_depress_t")     %>%
    rename_if_present("mh_p_cbcl__synd__rule_tscore",          "cbcl_scr_syn_rulebreak_t")    %>%
    rename_if_present("mh_p_cbcl__synd__rule_sum",             "cbcl_scr_syn_rulebreak_sum")  %>%
    rename_if_present("mh_p_cbcl__synd__int_tscore",           "cbcl_scr_syn_internal_t")     %>%
    rename_if_present("mh_p_cbcl__rule__cond_002",             "cbcl_q28_p")                  %>%
    rename_if_present("mh_p_cbcl__aggr__opp_003",              "cbcl_q23_p")                  %>%
    rename_if_present("su_y_tlfb__nic_ud",                     "su_tlfb_cal_scr_nic_days_yr") %>%
    rename_if_present("su_y_tlfb_ud",                          "tlfb_cal_scr_tot_su_ud")      %>%
    rename_if_present("fc_p_sag_005",                          "sag_grade_type")
}

p_demo            <- apply_renames(p_demo)
mh_p_ksads_bg     <- apply_renames(mh_p_ksads_bg)
led_l_seda_demo_d <- apply_renames(led_l_seda_demo_d)
ce_y_crpbi        <- apply_renames(ce_y_crpbi)
cbcl              <- apply_renames(cbcl)
su_y_tlfb         <- apply_renames(su_y_tlfb)
ce_p_sag          <- apply_renames(ce_p_sag)
su_y_sui          <- apply_renames(su_y_sui)
abcd_all          <- apply_renames(abcd_all)

cat("Renaming complete. Columns in p_demo:\n")
print(colnames(p_demo))

```

------------------------------------------------------------------------

## Step 5: Data Wrangling -- Build the Baseline Dataset

The ABCD variable `ab_g_stc__cohort_ethnrace__mhisp` (recoded as `race_ethnicity`) uses category 11 for "Multiracial Hispanic" and category 12 for "Multiracial non-Hispanic". We refine these by checking additional race questions:

- If they also selected Black (demo_race_a_p\_\_\_11 == 1), recode to 12 (Multiracial Black) This gives us more granular race groups for analysis.

- If they do not identify as black (demo_race_a_p\_\_\_11 != 1), recode to 13 (Other).

```{r recode-race}
# ============================================================
# RACE/ETHNICITY RECODING
# current coding for race_ethnicity
#   1 = Hispanic
#   2 = White
#   3 = Black
#   7 = Asian/Pacific Islander
#   11 = Multiracial + Hispanic
#   12 = Multiracial + non-Hispanic 
#   13 = Other 
# ============================================================
p_demo <- p_demo %>%
  mutate(race_ethnicity = case_when(
    race_ethnicity == 1                                  ~ 1, # hispanic
    race_ethnicity == 2                                  ~ 2, # white
    race_ethnicity == 3                                  ~ 3, # black
    race_ethnicity == 7                                  ~ 4, # asian/pacific islander
    race_ethnicity == 11                                 ~ 5, # multiracial + hispanic
    race_ethnicity == 12 & demo_race_a_p___11 == 1       ~ 6, # multiracial + black
    race_ethnicity == 12 & (demo_race_a_p___11 != 1
                            | is.na(demo_race_a_p___11)) ~ 7, # other
    race_ethnicity == 13                                 ~ 7, # other
    TRUE                                                 ~ race_ethnicity
  )) 

```

The ABCD data is longitudinal (multiple timepoints per child). For background/covariate variables, we only want BASELINE values (collected at the start of the study, when children were \~9-10 years old). We then merge all baseline datasets into one wide table using `reduce()` and `merge()`. The result is one row per participant with all covariates.

```{r subset-baseline}
# ============================================================
# SUBSET TO BASELINE TIMEPOINT
# ============================================================

# Filter each dataset to baseline only
p_demo_baseline            <- subset(p_demo,            eventname == "ses-00A")
mh_p_ksads_bg_baseline     <- subset(mh_p_ksads_bg,     eventname == "ses-00A")
led_l_seda_demo_d_baseline <- subset(led_l_seda_demo_d, eventname == "ses-00A")
ce_y_crpbi_baseline        <- subset(ce_y_crpbi,        eventname == "ses-00A")
cbcl_baseline              <- subset(cbcl,              eventname == "ses-00A")

# Merge all baseline datasets together by participant ID (src_subject_id).
# reduce() applies merge() repeatedly across the list.
# all = TRUE means keep all participants even if missing from one dataset.
background_df <- reduce(
  list(p_demo_baseline, mh_p_ksads_bg_baseline,
       led_l_seda_demo_d_baseline, ce_y_crpbi_baseline, cbcl_baseline),
  function(x, y) merge(x, y, by = "src_subject_id", all = TRUE)
)

# Select only the columns we need for analysis
# (keeps the dataset small and easier to work with)
filtered_df <- background_df %>%
  select(
    src_subject_id,
    site,                       # Data collection site
    demo_brthdat_v2,            # Age (at visit, i.e., baseline)
    demo_sex_v2,                # Sex (1=Male, 2=Female)
    race_ethnicity,             # Recoded race/ethnicity
    demo_prnt_income_v2,        # Household income bracket
    ledsch_seda_d_urban,        # Rural (1) vs. Non-rural (0)
    kbi_p_c_school_setting,     # School type (public/private/home)
    crpbi_caregiver1_y,         # Secondary caregiver support
    demo_prnt_marital_v2,       # Parental marital status
    cbcl_scr_syn_external_t,    # CBCL Externalizing T-score
    cbcl_scr_dsm5_depress_t,    # CBCL Depression T-score
    cbcl_scr_syn_rulebreak_t,   # CBCL Rule-Breaking T-score
   #cbcl_scr_syn_rulebreak_sum, # CBCL Rule-Breaking Sum 
    cbcl_q28_p,                 # CBCL item 28
    cbcl_q23_p                  # CBCL item 23
  )

# --- Parental Education (special handling) ---
# Education level can be updated at later timepoints, so we grab the
# MOST RECENT non-missing value per participant rather than baseline only.
parental_education <- p_demo %>%
  select(src_subject_id, demo_prnt_ed_v2_2yr_l) %>%
  drop_na(demo_prnt_ed_v2_2yr_l) %>%
  distinct(src_subject_id, .keep_all = TRUE)   # Keep one row per participant

nrow(filtered_df)
# Merge parental education back into the baseline dataset
filtered_df_baseline <- merge(filtered_df, parental_education, by = "src_subject_id", all.x = TRUE)

nrow(filtered_df_baseline)
```

Raw ABCD codes use many fine-grained numeric categories. We collapse these into broader brackets for cleaner analysis. Missing/refused/don't know responses are recoded to NA.

```{r recode-covariates}
# ============================================================
# RECODE CATEGORICAL VARIABLES INTO ANALYSIS-READY BRACKETS
# ============================================================

# --- Household Income ---
# New: 5 brackets (<50K, 50-75K, 75-100K, 100-200K, >200K)
filtered_df_baseline <- filtered_df_baseline %>%
  mutate(
    income_bracket = case_when(
      demo_prnt_income_v2 %in% 1:2   ~ 1,        # < $50,000
      demo_prnt_income_v2 == 3       ~ 2,        # $50,000 - $74,999
      demo_prnt_income_v2 == 4       ~ 3,        # $75,000 - $99,999
      demo_prnt_income_v2 == 5       ~ 4,        # $100,000 - $199,999
      demo_prnt_income_v2 == 6       ~ 5,        # >= $200,000
      demo_prnt_income_v2 == 777     ~ 777,      # Refused
      demo_prnt_income_v2 == 999     ~ 999,      # Don't know
      TRUE                           ~ NA_real_
    )
  ) %>%
  select(-demo_prnt_income_v2)   # Remove the original column (no longer needed)

# --- Parental Education ---
# New: 5 brackets from "Up to high school (no diploma)" to "Graduate degree or professional"
filtered_df_baseline <- filtered_df_baseline %>%
  mutate(
    parental_ed_bracket = case_when(
      demo_prnt_ed_v2_2yr_l == 777                 ~ NA_real_, # Refused
      demo_prnt_ed_v2_2yr_l == 999                 ~ NA_real_, # Don't know
      TRUE                                         ~ demo_prnt_ed_v2_2yr_l
    )
  )

# --- Parental Marital Status ---
# Original: 1-6 scale + 777
# New: 3 groups (Married/Partner, Separated/Divorced/Widowed, Never Married)
filtered_df_baseline <- filtered_df_baseline %>%
  mutate(
    marital_status_bracket = case_when(
      demo_prnt_marital_v2 %in% c(1, 6)    ~ 1,        # Married or living with partner
      demo_prnt_marital_v2 %in% c(2, 3, 4) ~ 2,        # Widowed / Divorced / Separated
      demo_prnt_marital_v2 == 5            ~ 3,        # Never married
      demo_prnt_marital_v2 == 777          ~ NA_real_, # Refused
      TRUE                                 ~ NA_real_
    )
  ) %>%
  select(-demo_prnt_marital_v2)   # Remove original column

# Convert sex, urban, marital status and caregiver to factor (categorical) type
filtered_df_baseline$demo_sex_v2 <- as.factor(filtered_df_baseline$demo_sex_v2)
filtered_df_baseline$ledsch_seda_d_urban <- as.factor(filtered_df_baseline$ledsch_seda_d_urban)
filtered_df_baseline$crpbi_caregiver1_y  <- as.factor(filtered_df_baseline$crpbi_caregiver1_y)
filtered_df_baseline$marital_status_bracket <- as.factor(filtered_df_baseline$marital_status_bracket)
```

Suspension data (kbi_p_c_det_susp_l) is in long-format across timepoints.

GOAL: Create a binary variable "ed_det_susp_baseline_3" that equals:

\- 1 = child was suspended at LEAST ONCE by the 3-year follow-up

\- 0 = child was never suspended through the 3-year follow-up

We also create "ed_det_susp_count" = total number of timepoints with a recorded suspension (0, 1, or 3 out of the first three follow-ups).

::: callout-note
Baseline suspension data is all missing in this dataset, so we only use year 1, 2 and 3 follow-up timepoints.
:::

```{r build-suspension-outcome}
# ============================================================
# BUILD THE SUSPENSION OUTCOME VARIABLE
# ============================================================

# Recode raw values: 1=Yes, 0=No -> 0, anything else -> NA
mh_p_ksads_bg <- mh_p_ksads_bg %>%
  mutate(
    administered = TRUE,  # every row in long format means the item was administered that wave
    kbi_p_c_det_susp_l = case_when(
      kbi_p_c_det_susp_l == 999   ~ NA_real_,  # not sure
      kbi_p_c_det_susp_l == 777   ~ NA_real_,  # declined to answer
      TRUE                        ~ kbi_p_c_det_susp_l
    )
  )

# Pivot suspension data to wide format (one column per timepoint)
susp <- mh_p_ksads_bg[, c("src_subject_id", "eventname", "kbi_p_c_det_susp_l", "administered")]
susp_wide <- susp %>%
  pivot_wider(
    id_cols     = src_subject_id,
    names_from  = eventname,
    values_from = c(kbi_p_c_det_susp_l, administered), 
    names_glue  = "{eventname}_{.value}"
  ) %>%
  rename_with(
    .fn   = ~ str_remove(., "_kbi_p_c_det_susp_l"),  # strip the value-name suffix from the disciplinary action columns
    .cols = ends_with("_kbi_p_c_det_susp_l")
  ) %>% 
  rename_with(
    .fn   = ~ paste0("ed_", .),   # Prefix all timepoint columns with "ed_"
    .cols = -src_subject_id
  )

# Create binary "ever suspended" indicator using years 1, 2 and 3 follow-up
# 1 = suspended at year 1 OR year 2 OR year 3; 0 = not suspended at any
# NA is only assigned when the participants has NA across all three waves
# this treats "missing at wave 1 and 2, but 0 at wave 3" as a confirmed 0 for the whole period
susp_wide <- susp_wide %>%
  mutate(
    n_nonmissing = rowSums(!is.na(select(., `ed_ses-01A`, `ed_ses-02A`, `ed_ses-03A`))),
    n_yes        = rowSums(select(., `ed_ses-01A`, `ed_ses-02A`, `ed_ses-03A`) == 1, na.rm = TRUE),
    ed_det_susp_baseline_3 = case_when(
      n_yes > 0        ~ 1,
      n_nonmissing == 0 ~ NA_real_,
      TRUE              ~ 0
    )
  ) %>%
  select(-n_nonmissing, -n_yes)

# old code by Siqi
#susp_wide <- susp_wide %>%
  #mutate(
    #ed_det_susp_baseline_3 = if_else(
      #`ed_ses-01A` == 1 |
      #`ed_ses-02A` == 1 | 
      #`ed_ses-03A` == 1,
      #1, 0
    #)
  #)

# Create count variable: how many timepoints had a suspension (0, 1, or 2)
susp_wide <- susp_wide %>%
  mutate(
    ed_det_susp_count = rowSums(
      select(., "ed_ses-01A", "ed_ses-02A") == 1,
      na.rm = TRUE
    )
  )

# Merge suspension data onto the baseline covariate dataset
filtered_df_baseline_ed <- left_join(filtered_df_baseline, susp_wide, by = "src_subject_id")
```

Convert numeric codes to readable category labels. This makes tables and plots automatically show human-readable text instead of numbers like "1", "2", "3".

```{r recode-labels}
# ============================================================
# APPLY FACTOR LABELS FOR TABLE DISPLAY
# ============================================================

filtered_df_baseline_recoded <- filtered_df_baseline_ed %>%
  mutate(
    race_ethnicity = factor(
      race_ethnicity,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c("Hispanic", "White", "Black", "Asian and Pacific Islander",
                 "Multiracial Hispanic", "Multiracial Black", "Other")
    ),
    demo_sex_v2 = factor(
      demo_sex_v2,
      levels = c(1, 2),
      labels = c("Male", "Female")
    ),
    ledsch_seda_d_urban = factor(
      ledsch_seda_d_urban,
      levels = c(1, 2, 3),
      labels = c("Urban", "Urban clusters", "Rural")
    ),
    # School type: several codes collapse into "Other"
    kbi_p_c_school_setting = factor(
      kbi_p_c_school_setting,
      levels = c(2, 3, 6, 1, 4, 5, 7, 8, 9),
      labels = c("Regular Public School", "Regular Private School", "Home School",
                 "Other", "Other", "Other", "Other", "Other", "Other")
    ),
    crpbi_caregiver1_y = factor(
      crpbi_caregiver1_y,
      levels = c(0, 1),
      labels = c("No", "Yes")
    ),
    income_bracket = factor(
      income_bracket,
      levels = c(1, 2, 3, 4, 5, 777, 999),
      labels = c("<$50K", "$50K-$75K", "$75K-$100K", "$100K-$200K", ">$200K", "Declined", "Don't know")
    ),
    parental_ed_bracket = factor(
      parental_ed_bracket,
      levels = c(1, 2, 3, 4, 5),
      labels = c("Up to high school (no diploma)", "High school diploma/GED", "Some College", "Bachelor's degree", "Graduate or professional degree")
    ),
    marital_status_bracket = factor(
      marital_status_bracket,
      levels = c(1, 2, 3),
      labels = c("Married/Living Together", "Separated/Divorced/Widowed", "Never Married")
    ),
    ed_det_susp_baseline_3 = factor(
      ed_det_susp_baseline_3,
      levels = c(0, 1),
      labels = c("No", "Yes")
    )
  )

# Apply human-readable variable labels for Table 1 display
# (These labels appear as row headers in the summary table)
label(filtered_df_baseline_recoded$demo_brthdat_v2)           <- "Age (at Baseline)"
label(filtered_df_baseline_recoded$site)                      <- "Data collection site"
label(filtered_df_baseline_recoded$demo_sex_v2)               <- "Sex"
label(filtered_df_baseline_recoded$race_ethnicity)            <- "Race/Ethnicity"
label(filtered_df_baseline_recoded$ledsch_seda_d_urban)       <- "Urban vs. Rural"
label(filtered_df_baseline_recoded$kbi_p_c_school_setting)    <- "Type of School"
label(filtered_df_baseline_recoded$crpbi_caregiver1_y)        <- "2nd Caregiver Support"
label(filtered_df_baseline_recoded$cbcl_scr_syn_external_t)   <- "Externalizing T-score (CBCL)"
label(filtered_df_baseline_recoded$income_bracket)            <- "Household Income"
label(filtered_df_baseline_recoded$parental_ed_bracket)       <- "Parental Education Bracket"
label(filtered_df_baseline_recoded$marital_status_bracket)    <- "Parental Marital Status"
label(filtered_df_baseline_recoded$cbcl_scr_dsm5_depress_t)   <- "Depression T-score (CBCL)"
label(filtered_df_baseline_recoded$cbcl_scr_syn_rulebreak_t)  <- "Rule Breaking T-score (CBCL)"
label(filtered_df_baseline_recoded$ed_det_susp_baseline_3)    <- "Ever received ED (Baseline to 3yr)"

# Print a quick data structure summary to verify the dataset looks correct
str(filtered_df_baseline_recoded)
nrow(filtered_df_baseline_recoded)
```

------------------------------------------------------------------------

## Step 6: Descriptive Statistics by Race Subgroup

`describe()` from the Hmisc package prints detailed summaries (counts, means, distributions) for all variables. We run this on the raw (un-labeled) dataset. First for the White participants.

```{r describe-by-race}
# ============================================================
# QUICK DESCRIPTIVE SUMMARIES FOR RACE SUBGROUPS
# ============================================================

# Subset for White participants (race code 2)
filtered_df_baseline_white     <- subset(filtered_df_baseline, race_ethnicity == 2)
# Subset for Black and Hispanic participants (codes 1, 3, 5, 6)
filtered_df_baseline_black_his <- subset(filtered_df_baseline, race_ethnicity %in% c(1, 3, 5, 6))

describe(filtered_df_baseline_white)
```

Then for the Black and Hispanic participants.

```{r describe-black-hispanic}
describe(filtered_df_baseline_black_his)
```

------------------------------------------------------------------------

## Step 7: Build Longitudinal Outcome Datasets

Create a wide-format dataset for nicotine/tobacco use. Also create `"ever_used_tobacco" = 1` if ANY nonzero tobacco use was reported across all timepoints.

```{r tobacco-wide}
# ============================================================
# TOBACCO USE OUTCOME (WIDE FORMAT)
# ============================================================
tabacco <- su_y_tlfb[, c("src_subject_id", "eventname", "su_tlfb_cal_scr_nic_days_yr")]
tabacco_wide <- tabacco %>%
  pivot_wider(
    id_cols     = src_subject_id,
    names_from  = eventname,
    values_from = su_tlfb_cal_scr_nic_days_yr
  ) %>%
  rename_with(
    .fn   = ~ paste0("tobacco_", .),
    .cols = -src_subject_id
  )

# Create binary "ever used tobacco" indicator
tabacco_wide <- tabacco_wide %>%
  mutate(
    ever_used_tobacco = ifelse(
      rowSums(select(., starts_with("tobacco_")), na.rm = TRUE) > 0, 1, 0
    )
  )

describe(tabacco_wide)
```

Quick check on the distribution of our key exposure variable

```{r check-suspension-outcome}
filtered_df_baseline_ed$ed_det_susp_baseline_3 <- as.factor(
  filtered_df_baseline_ed$ed_det_susp_baseline_3
)
describe(filtered_df_baseline_ed$ed_det_susp_baseline_3)
```

Check suspension rates separately for White and Black/Hispanic participants

```{r describe-suspension-by-race}
filtered_df_baseline_ed_white     <- subset(filtered_df_baseline_ed, race_ethnicity == 2)
filtered_df_baseline_ed_black_his <- subset(filtered_df_baseline_ed, race_ethnicity %in% c(1, 3, 5, 6))

describe(filtered_df_baseline_ed_white$ed_det_susp_baseline_3)
describe(filtered_df_baseline_ed_black_his$ed_det_susp_baseline_3)
```

------------------------------------------------------------------------

## Step 8: Missing Data Mechanisms

GOAL: Assess the extent and patterns of missing data.

```{r miss-baseline}
filtered_df_baseline %>%
  gg_miss_var(show_pct = TRUE) + labs(title = "Proportion of missing data at baseline") + theme_brand()

df_plot <- filtered_df_baseline %>% select(-src_subject_id)

aggr_plot <- aggr(df_plot, plot = FALSE)  # compute without plotting

plot(aggr_plot, 
     numbers = TRUE, 
     prop = FALSE,
     combined = TRUE,
     labels = names(df_plot),
     cex.axis = 0.4,
     las = 2)   # rotate labels vertical
```

Assess the proportion missing on detentions/suspension over time.

```{r miss-suspenison, warning=FALSE, message=FALSE, fig.width=10, fig.height=8}
# make variable titles clearer
ed_plot <- filtered_df_baseline_recoded %>%
  mutate(across(
    c(`ed_ses-00A_administered`, `ed_ses-01A_administered`, `ed_ses-02A_administered`,
      `ed_ses-03A_administered`, `ed_ses-04A_administered`, `ed_ses-05A_administered`,
      `ed_ses-06A_administered`, `ed_ses-07A_administered`),
    ~ replace_na(.x, FALSE)
  )) %>%
  mutate(
    `ed_ses-01A` = if_else(!`ed_ses-01A_administered`, -999, `ed_ses-01A`),
    `ed_ses-02A` = if_else(!`ed_ses-02A_administered`, -999, `ed_ses-02A`),
    `ed_ses-03A` = if_else(!`ed_ses-03A_administered`, -999, `ed_ses-03A`)
  ) %>%
  mutate(
    `Detention Baseline` = factor(`ed_ses-00A`, labels = c("No", "Yes")),
    `Detention 1YR` = factor(`ed_ses-01A`, levels = c(-999, 0, 1), labels = c("Not administered", "No", "Yes")),
    `Detention 2YR` = factor(`ed_ses-02A`, levels = c(-999, 0, 1), labels = c("Not administered", "No", "Yes")),
    `Detention 3YR` = factor(`ed_ses-03A`, levels = c(-999, 0, 1), labels = c("Not administered", "No", "Yes")),
    `Detention 4YR` = factor(`ed_ses-04A`, labels = c("No", "Yes")),
    `Detention 5YR` = factor(`ed_ses-05A`, labels = c("No", "Yes")),
    `Detention 6YR` = factor(`ed_ses-06A`, labels = c("No", "Yes")),
    `Detention 7YR` = factor(`ed_ses-07A`, labels = c("No", "Yes"))
  )

ed_plot %>%
  select(`Detention Baseline`, `Detention 1YR`, `Detention 2YR`, `Detention 3YR`, `Detention 4YR`, `Detention 5YR`, `Detention 6YR`, `Detention 7YR`) %>%
  gg_miss_var(show_pct = TRUE) + labs(title = "Proportion of missing data for detention/suspension at each wave") + theme_brand()

ed_plot %>%
  select(`Detention Baseline`, `Detention 1YR`, `Detention 2YR`, `Detention 3YR`, `Detention 4YR`, `Detention 5YR`, `Detention 6YR`, `Detention 7YR`) %>%
  vis_miss(cluster = TRUE, show_perc = FALSE) +
  scale_fill_manual(
    values = theme_colors[c(5, 19)],
    labels = c("Present", "Missing"),
    name = ""
  ) +
  theme_brand()
```

```{r miss-suspenison-patterns, fig.width=10, fig.height=8, message=FALSE}
ed_plot %>%
  select(`Detention Baseline`, `Detention 1YR`, `Detention 2YR`, `Detention 3YR`, `Detention 4YR`, `Detention 5YR`, `Detention 6YR`, `Detention 7YR`, race_ethnicity) %>%
  gg_miss_var(show_pct = TRUE, facet = race_ethnicity) + labs(title = "Proportion of missing data for detention/suspension at each wave by ethnoracial identity") + theme_brand()

ed_plot %>% 
  select(`Detention Baseline`, `Detention 1YR`, `Detention 2YR`, `Detention 3YR`, `Detention 4YR`, `Detention 5YR`, `Detention 6YR`, `Detention 7YR`, race_ethnicity) %>%
  gg_miss_fct(race_ethnicity) +
  scale_fill_gradientn(
    guide = "colorbar",
    name = "% Miss",
    colors = theme_colors[c(1:10)]
  ) +
  ylab("Variable") +
  xlab("Ethno-racial identity") +
  theme_brand()
```

```{r miss-suspenison-by-demo, message=FALSE}
ed_plot %>% 
  select(`Detention Baseline`, `Detention 1YR`, `Detention 2YR`, `Detention 3YR`, `Detention 4YR`, `Detention 5YR`, `Detention 6YR`, `Detention 7YR`, demo_sex_v2) %>%
  gg_miss_fct(demo_sex_v2) +
  scale_fill_gradientn(
    guide = "colorbar",
    name = "% Miss",
    colors = theme_colors[c(1:10)]
  ) +
  ylab("Variable") +
  xlab("Sex") +
  theme_brand()

ed_plot %>% 
  select(`Detention Baseline`, `Detention 1YR`, `Detention 2YR`, `Detention 3YR`, `Detention 4YR`, `Detention 5YR`, `Detention 6YR`, `Detention 7YR`, income_bracket) %>%
  gg_miss_fct(income_bracket) +
  scale_fill_gradientn(
    guide = "colorbar",
    name = "% Miss",
    colors = theme_colors[c(1:10)]
  ) +
  ylab("Variable") +
  xlab("Parental income") +
  theme_brand()
```

------------------------------------------------------------------------

## Step 9: Propensity Score Matching -- White vs. Black+Hispanic

GOAL: Create a matched sample where White and minoritized children are comparable on confounding variables (SES, school type, etc.) so that any difference in suspension rates reflects race, not background differences.

TREATMENT GROUP: White (race_ethnicity_recode = 1)

CONTROL GROUP: Black + Hispanic (race_ethnicity_recode = 0)

MATCHING VARIABLES: marital status, parental education, income, secondary caregiver, school setting, urban/rural, sex, age, CBCL externalizing T-score (behavior problems at baseline)

```{r psm-wb-his-prep}
# ============================================================
# PROPENSITY SCORE MATCHING: WHITE vs. BLACK + HISPANIC
# ============================================================

# Recode race into binary treatment indicator
filtered_df_baseline_ed_recode <- filtered_df_baseline_ed %>%
  mutate(
    race_ethnicity_recode = case_when(
      race_ethnicity == 2                ~ 1,   # White = treated
      race_ethnicity %in% c(1, 3, 5, 6)  ~ 0,   # Black/Hispanic = control
      TRUE                               ~ NA_real_
    )
  ) %>%
  filter(!is.na(race_ethnicity_recode))   # Drop Asian and Other groups

# Select only variables needed for matching, drop rows with any missing values
prop_subset_data <- filtered_df_baseline_ed_recode %>%
  select(
    src_subject_id,
    site,                    # Data collection site
    demo_brthdat_v2,         # Age
    demo_sex_v2,             # Sex
    ledsch_seda_d_urban,     # Urban/rural
    kbi_p_c_school_setting,  # School type
    crpbi_caregiver1_y,      # Secondary caregiver support
    income_bracket,          # Household income
    parental_ed_bracket,     # Parental education (continuous)
    marital_status_bracket,  # Marital status
    race_ethnicity_recode,   # Binary race indicator (what we match on)
    race_ethnicity,          # Original race code (kept for reference)
    cbcl_scr_syn_external_t  # CBCL externalizing T-score (behavior control)
  ) %>%
  na.omit()

cat("White participants (treated):",   sum(prop_subset_data$race_ethnicity_recode == 1), "\n")
cat("Black+Hispanic (control):",       sum(prop_subset_data$race_ethnicity_recode == 0), "\n")
```

Define the matching formula (variables to match on) and perform nearest-neighbor 1:1 matching with a caliper of 0.2. `caliper = 0.2` means matches must be within 0.2 standard deviations of propensity score -- prevents poor/distant matches

```{r psm-wb-his-match}
matching_formula <- race_ethnicity_recode ~
  marital_status_bracket + parental_ed_bracket + income_bracket +
  crpbi_caregiver1_y + kbi_p_c_school_setting + ledsch_seda_d_urban +
  demo_sex_v2 + demo_brthdat_v2 + cbcl_scr_syn_external_t

matched <- matchit(
  matching_formula,
  data    = prop_subset_data,
  method  = "nearest",
  ratio   = 1,
  caliper = 0.2
)
summary(matched)
```

Check the propensity score balance.

BEFORE matching: the two groups may have very different propensity score distributions (curves don't overlap much).

AFTER matching: the curves should overlap closely, showing the groups are now comparable on measured confounders.

```{r psm-wb-his-plot}
# ============================================================
# KERNEL DENSITY PLOTS: CHECK PROPENSITY SCORE BALANCE
# ============================================================

# Calculate propensity scores for all participants (before matching)
prop_subset_data <- prop_subset_data %>%
  mutate(
    propensity_score = predict(
      glm(matching_formula, family = binomial, data = .),
      type = "response"
    )
  )

matched_data <- match.data(matched)   # Extract the matched sample

# Plot BEFORE matching
before_matching_plot <- ggplot(
  prop_subset_data,
  aes(x = propensity_score, fill = as.factor(race_ethnicity_recode))
) +
  geom_density(alpha = 0.5) +
  labs(title = "Kernel Density Plot Before Matching",
       x = "Propensity Score", y = "Density", fill = "Group") +
  scale_fill_manual(values = c("#CC4E6A", "#3C4E97"),
                    labels = c("Minoritized (Control)", "White (Treated)")) +
  theme_brand()

ggsave("output/kernel_density_before_matching.png",
       before_matching_plot, width = 8, height = 6)

# Plot AFTER matching
after_matching_plot <- ggplot(
  matched_data,
  aes(x = distance, fill = as.factor(race_ethnicity_recode))
) +
  geom_density(alpha = 0.5) +
  labs(title = "Kernel Density Plot After Matching",
       x = "Propensity Score", y = "Density", fill = "Group") +
  scale_fill_manual(values = c("#CC4E6A", "#3C4E97"),
                    labels = c("Minoritized (Control)", "White (Treated)")) +
  theme_brand()

ggsave("output/kernel_density_after_matching.png",
       after_matching_plot, width = 8, height = 6)
```

```{r, message=FALSE, warning=FALSE}
# Plot balance statistics
love_plot_after <- cobalt::love.plot(matched, binary = "std", sample.names = c("Unmatched", "Matched")) + 
  scale_y_discrete(labels = c("distance" = "Distance", "marital_status_bracket_1" = "Married/Living Together", "marital_status_bracket_2" = "Separated/Divorced/Widowed", "marital_status_bracket_3" = "Never Married", "parental_ed_bracket" = "Partner Educational Attainment", "income_bracket" = "Income Bracket", "crpbi_caregiver1_y" = "Additional Caregiver", "kbi_p_c_school_setting" = "School Setting", "ledsch_seda_d_urban_1" = "Urban Locale", "ledsch_seda_d_urban_2" = "Urbanised Clusters", "ledsch_seda_d_urban_3" = "Rural Locale", "demo_sex_v2_2" = "Sex", "demo_brthdat_v2" = "Age", "cbcl_scr_syn_external_t" = "Externalizing T-score")) +
  theme_brand()

class(matched_data$parental_ed_bracket)
class(matched_data$ledsch_seda_d_urban)

love_plot_after

ggsave("output/love.plot_after_matching.png",
       love_plot_after, width = 8, height = 6)
```

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=12}
cowplot::plot_grid(before_matching_plot, after_matching_plot, nrow =2)
```

------------------------------------------------------------------------

## Step 10: Build Longitudinal Outcome Wide Datasets

We need wide-format datasets for four outcomes:

1\. Depression T-scores (cbcl_scr_dsm5_depress_t) -- continuous

2\. Substance use (tlfb_cal_scr_tot_su_ud) -- continuous

3\. Academic grades (grade_gpa) -- converted letter -\> GPA

4\. Internalizing T-scores (cbcl_scr_syn_internal_t) -- continuous

Each uses the `pivot_wide()` function defined at the top of this document.

```{r build-outcome-datasets}
# ============================================================
# PIVOT OUTCOME VARIABLES TO WIDE FORMAT
# ============================================================

# 1. Depression T-scores across all timepoints
depression      <- cbcl[, c("src_subject_id", "eventname", "cbcl_scr_dsm5_depress_t")]
depression_wide <- pivot_wide(depression, "cbcl_scr_dsm5_depress_t")

# 2. Substance use across all timepoints
substance      <- su_y_tlfb[, c("src_subject_id", "eventname", "tlfb_cal_scr_tot_su_ud")]
substance_wide <- pivot_wide(substance, "tlfb_cal_scr_tot_su_ud")

# Binary "ever used any substance" indicator
substance_wide <- substance_wide %>%
  mutate(
    ever_used_substance = ifelse(
      rowSums(select(., starts_with("tlfb_cal_scr_tot_su_ud_")), na.rm = TRUE) > 0,
      1, 0
    )
  )

# 3. Academic grades: convert letter grade codes to GPA-equivalent numeric values
# Using the lower boundary for each grade range
grade <- ce_p_sag[, c("src_subject_id", "eventname", "sag_grade_type")]
grade <- grade %>%
  mutate(
    grade_gpa = case_when(
      sag_grade_type == 1  ~ 4.0,  # A+
      sag_grade_type == 2  ~ 3.7,  # A
      sag_grade_type == 3  ~ 3.5,  # A-
      sag_grade_type == 4  ~ 3.3,  # B+
      sag_grade_type == 5  ~ 3.0,  # B
      sag_grade_type == 6  ~ 2.7,  # B-
      sag_grade_type == 7  ~ 2.3,  # C+
      sag_grade_type == 8  ~ 2.0,  # C
      sag_grade_type == 9  ~ 1.7,  # C-
      sag_grade_type == 10 ~ 1.3,  # D+
      sag_grade_type == 11 ~ 1.0,  # D
      sag_grade_type == 12 ~ 0.0,  # F
      TRUE                 ~ NA_real_
    )
  )
grade_wide <- pivot_wide(grade, "grade_gpa")

# 4. Internalizing T-scores across all timepoints
internal_t      <- cbcl[, c("src_subject_id", "eventname", "cbcl_scr_syn_internal_t")]
internal_t_wide <- pivot_wide(internal_t, "cbcl_scr_syn_internal_t")
```

------------------------------------------------------------------------

## Step 11: Finalize Matched Dataset -- White vs. Black+Hispanic

Starting from the matched participants, we join:

\- Suspension data (susp_wide)

\- Depression scores (depression_wide)

\- Substance use (substance_wide)

\- Grades (grade_wide)

Then save as a CSV.

```{r finalize-matched-wb-his}
# ============================================================
# MERGE ALL OUTCOME DATASETS ONTO THE MATCHED SAMPLE
# ============================================================

merged_data_sus <- susp_wide %>%
  filter(src_subject_id %in% matched_data$src_subject_id) %>%
  left_join(matched_data, by = "src_subject_id")

merged_data_depression_sus <- depression_wide %>%
  filter(src_subject_id %in% merged_data_sus$src_subject_id) %>%
  left_join(merged_data_sus, by = "src_subject_id")

merged_data_depression_sus_su <- merged_data_depression_sus %>%
  left_join(substance_wide, by = "src_subject_id")

merged_data_depression_sus_su_grade <- merged_data_depression_sus_su %>%
  left_join(grade_wide, by = "src_subject_id")

# Remove matching-specific columns before saving
matched_data_raw <- merged_data_depression_sus_su_grade %>%
  select(-propensity_score, -distance, -weights, -subclass)

write.csv(matched_data_raw,
          "data/matched_data_raw.csv",
          row.names = FALSE)
cat("Saved: output/matched_data_raw.csv\n")

# ============================================================
# PRE-MATCH DATASET (full unmatched sample)
# Used for sensitivity analyses comparing matched vs. unmatched results
# ============================================================

prematch_depression_sus <- depression_wide %>%
  filter(src_subject_id %in% filtered_df_baseline_ed_recode$src_subject_id) %>%
  left_join(filtered_df_baseline_ed_recode, by = "src_subject_id")

prematch_depression_sus_su <- prematch_depression_sus %>%
  left_join(substance_wide, by = "src_subject_id")

prematch_depression_sus_su_grade <- prematch_depression_sus_su %>%
  left_join(grade_wide, by = "src_subject_id")

prematch_data_raw <- prematch_depression_sus_su_grade %>%
  left_join(internal_t_wide, by = "src_subject_id")

write.csv(prematch_data_raw,
          "data/prematch_data_raw.csv",
          row.names = FALSE)
cat("Saved: prematch_data_raw.csv\n")
```

------------------------------------------------------------------------

## Step 12: Propensity Score Matching -- White vs. Black

Same approach as Step 9, but narrowing to White vs. Black children. `ratio = 4` means up to 4 White children matched per Black child (used because there are many more White participants).

```{r psm-wb-prep}
# ============================================================
# PROPENSITY SCORE MATCHING: WHITE vs. BLACK ONLY
# ============================================================

WB_baseline_ed_recode <- filtered_df_baseline_ed %>%
  mutate(
    race_ethnicity_recode = case_when(
      race_ethnicity == 2           ~ 0,   # White = control (reference)
      race_ethnicity %in% c(3, 6)   ~ 1,   # Black / Multiracial Black = treated
      TRUE                          ~ NA_real_
    )
  ) %>%
  filter(!is.na(race_ethnicity_recode))

WB_prop_subset_data <- WB_baseline_ed_recode %>%
  select(
    src_subject_id, site, demo_brthdat_v2, demo_sex_v2, ledsch_seda_d_urban,
    kbi_p_c_school_setting, crpbi_caregiver1_y, income_bracket,
    parental_ed_bracket, marital_status_bracket,
    race_ethnicity_recode, race_ethnicity,
    cbcl_scr_syn_external_t, cbcl_scr_syn_rulebreak_t #cbcl_scr_syn_rulebreak_sum
  ) %>%
  na.omit()

cat("Black participants (treated):", sum(WB_prop_subset_data$race_ethnicity_recode == 1), "\n")
cat("White participants (control):", sum(WB_prop_subset_data$race_ethnicity_recode == 0), "\n")

matching_formula <- race_ethnicity_recode ~
  marital_status_bracket + parental_ed_bracket + income_bracket +
  crpbi_caregiver1_y + kbi_p_c_school_setting + ledsch_seda_d_urban +
  demo_sex_v2 + demo_brthdat_v2 + cbcl_scr_syn_external_t

WB_matched <- matchit(
  matching_formula,
  data    = WB_prop_subset_data,
  method  = "nearest",
  ratio   = 4,
  caliper = 0.2
)
summary(WB_matched)
```

```{r psm-wb-plot}
WB_prop_subset_data <- WB_prop_subset_data %>%
  mutate(
    propensity_score = predict(
      glm(matching_formula, family = binomial, data = .),
      type = "response"
    )
  )

WB_matched_data <- match.data(WB_matched)

WB_before_matching_plot <- ggplot(
  WB_prop_subset_data,
  aes(x = propensity_score, fill = as.factor(race_ethnicity_recode))
) +
  geom_density(alpha = 0.5) +
  labs(title = "Kernel Density Plot Before Matching (White vs. Black)",
       x = "Propensity Score", y = "Density", fill = "Group") +
  scale_fill_manual(values = c("#CC4E6A", "#3C4E97"),
                    labels = c("White", "Black / Multiracial Black")) +
  theme_brand()

ggsave("output/WB_kernel_density_before_matching.png",
       WB_before_matching_plot, width = 8, height = 6)

WB_after_matching_plot <- ggplot(
  WB_matched_data,
  aes(x = distance, fill = as.factor(race_ethnicity_recode))
) +
  geom_density(alpha = 0.5) +
  labs(title = "Kernel Density Plot After Matching (White vs. Black)",
       x = "Propensity Score", y = "Density", fill = "Group") +
  scale_fill_manual(values = c("#CC4E6A", "#3C4E97"),
                    labels = c("White", "Black / Multiracial Black")) +
  theme_brand()

ggsave("output/WB_kernel_density_after_matching.png",
       WB_after_matching_plot, width = 8, height = 6)
```

```{r, message=FALSE, warning=FALSE}
# Plot balance statistics
WB_love_plot <- cobalt::love.plot(WB_matched, binary = "std", sample.names = c("Unmatched", "Matched")) + 
    scale_y_discrete(labels = c("distance" = "Distance", "marital_status_bracket_1" = "Married/Living Together", "marital_status_bracket_2" = "Separated/Divorced/Widowed", "marital_status_bracket_3" = "Never Married", "parental_ed_bracket" = "Partner Educational Attainment", "income_bracket" = "Income Bracket", "crpbi_caregiver1_y" = "Additional Caregiver", "kbi_p_c_school_setting" = "School Setting", "ledsch_seda_d_urban_1" = "Urban Locale", "ledsch_seda_d_urban_2" = "Urbanised Clusters", "ledsch_seda_d_urban_3" = "Rural Locale", "demo_sex_v2_2" = "Sex", "demo_brthdat_v2" = "Age", "cbcl_scr_syn_external_t" = "Externalizing T-score")) +
  theme_brand()

ggsave("output/WB_love.plot_after_matching.png",
       WB_love_plot, width = 8, height = 6)
```

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=14}
cowplot::plot_grid(WB_before_matching_plot, WB_after_matching_plot, WB_love_plot, nrow = 3)

WB_PSM <- cowplot::plot_grid(WB_before_matching_plot, WB_after_matching_plot, WB_love_plot, nrow = 3)

ggsave("output/WB_PSM.png",
       WB_PSM, width = 6, height = 9)
```

Merge and save matched dataset for White vs. Black comparison.

```{r finalize-matched-wb}
WB_merged_data_sus <- susp_wide %>%
  filter(src_subject_id %in% WB_matched_data$src_subject_id) %>%
  left_join(WB_matched_data, by = "src_subject_id")

WB_merged_data_depression_sus <- depression_wide %>%
  filter(src_subject_id %in% WB_merged_data_sus$src_subject_id) %>%
  left_join(WB_merged_data_sus, by = "src_subject_id")

WB_merged_data_depression_sus_su <- WB_merged_data_depression_sus %>%
  left_join(substance_wide, by = "src_subject_id")

WB_merged_data_depression_sus_su_grade <- WB_merged_data_depression_sus_su %>%
  left_join(grade_wide, by = "src_subject_id")

WB_matched_data_raw <- WB_merged_data_depression_sus_su_grade %>%
  left_join(internal_t_wide, by = "src_subject_id") %>%
  select(-propensity_score, -distance, -weights, -subclass)

write.csv(WB_matched_data_raw,
          "data/WB_matched_data_raw.csv",
          row.names = FALSE)
cat("Saved: WB_matched_data_raw.csv\n")

# Pre-match dataset for White vs. Black
WB_prematch_data_raw <- depression_wide %>%
  filter(src_subject_id %in% WB_baseline_ed_recode$src_subject_id) %>%
  left_join(WB_baseline_ed_recode, by = "src_subject_id") %>%
  left_join(substance_wide,  by = "src_subject_id") %>%
  left_join(grade_wide,      by = "src_subject_id") %>%
  left_join(tabacco_wide,    by = "src_subject_id") %>%
  left_join(internal_t_wide, by = "src_subject_id")

write.csv(WB_prematch_data_raw,
          "data/WB_prematch_data_raw.csv",
          row.names = FALSE)
cat("Saved: WB_prematch_data_raw.csv\n")
```


------------------------------------------------------------------------

## Step 13: Propensity Score Matching -- White vs. Hispanic

Same approach as Steps 9 and 12, now for White vs. Hispanic children.

```{r psm-wh-prep}
# ============================================================
# PROPENSITY SCORE MATCHING: WHITE vs. HISPANIC ONLY
# ============================================================

WH_baseline_ed_recode <- filtered_df_baseline_ed %>%
  mutate(
    race_ethnicity_recode = case_when(
      race_ethnicity == 2           ~ 0,   # White = control
      race_ethnicity %in% c(1, 5)   ~ 1,   # Hispanic / Multiracial Hispanic = treated
      TRUE                          ~ NA_real_
    )
  ) %>%
  filter(!is.na(race_ethnicity_recode))

WH_prop_subset_data <- WH_baseline_ed_recode %>%
  select(
    src_subject_id, site, demo_brthdat_v2, demo_sex_v2, ledsch_seda_d_urban,
    kbi_p_c_school_setting, crpbi_caregiver1_y, income_bracket,
    parental_ed_bracket, marital_status_bracket,
    race_ethnicity_recode, race_ethnicity,
    cbcl_scr_syn_external_t, cbcl_scr_syn_rulebreak_t #cbcl_scr_syn_rulebreak_sum
  ) %>%
  na.omit()

cat("White participants (control):",    sum(WH_prop_subset_data$race_ethnicity_recode == 0), "\n")
cat("Hispanic participants (treated):", sum(WH_prop_subset_data$race_ethnicity_recode == 1), "\n")

matching_formula <- race_ethnicity_recode ~
  marital_status_bracket + parental_ed_bracket + income_bracket +
  crpbi_caregiver1_y + kbi_p_c_school_setting + ledsch_seda_d_urban +
  demo_sex_v2 + demo_brthdat_v2 + cbcl_scr_syn_external_t

WH_matched <- matchit(
  matching_formula,
  data    = WH_prop_subset_data,
  method  = "nearest",
  ratio   = 4,
  caliper = 0.2
)
summary(WH_matched)
```

```{r psm-wh-plot}
WH_prop_subset_data <- WH_prop_subset_data %>%
  mutate(
    propensity_score = predict(
      glm(matching_formula, family = binomial, data = .),
      type = "response"
    )
  )

WH_matched_data <- match.data(WH_matched)

WH_before_matching_plot <- ggplot(
  WH_prop_subset_data,
  aes(x = propensity_score, fill = as.factor(race_ethnicity_recode))
) +
  geom_density(alpha = 0.5) +
  labs(title = "Kernel Density Plot Before Matching (White vs. Hispanic)",
       x = "Propensity Score", y = "Density", fill = "Group") +
  scale_fill_manual(values = c("#CC4E6A", "#3C4E97"), labels = c("White", "Latine")) +
  theme_brand()

ggsave("output/WH_kernel_density_before_matching.png",
       WH_before_matching_plot, width = 8, height = 6)

WH_after_matching_plot <- ggplot(
  WH_matched_data,
  aes(x = distance, fill = as.factor(race_ethnicity_recode))
) +
  geom_density(alpha = 0.5) +
  labs(title = "Kernel Density Plot After Matching (White vs. Hispanic)",
       x = "Propensity Score", y = "Density", fill = "Group") +
  scale_fill_manual(values = c("#CC4E6A", "#3C4E97"), labels = c("White", "Latine")) +
  theme_brand()

ggsave("output/WH_kernel_density_after_matching.png",
       WH_after_matching_plot, width = 8, height = 6)
```

```{r, message=FALSE, warning=FALSE}
# Plot balance statistics
WH_love_plot <- cobalt::love.plot(WH_matched, binary = "std", sample.names = c("Unmatched", "Matched")) + 
    scale_y_discrete(labels = c("distance" = "Distance", "marital_status_bracket_1" = "Married/Living Together", "marital_status_bracket_2" = "Separated/Divorced/Widowed", "marital_status_bracket_3" = "Never Married", "parental_ed_bracket" = "Partner Educational Attainment", "income_bracket" = "Income Bracket", "crpbi_caregiver1_y" = "Additional Caregiver", "kbi_p_c_school_setting" = "School Setting", "ledsch_seda_d_urban_1" = "Urban Locale", "ledsch_seda_d_urban_2" = "Urbanised Clusters", "ledsch_seda_d_urban_3" = "Rural Locale", "demo_sex_v2_2" = "Sex", "demo_brthdat_v2" = "Age", "cbcl_scr_syn_external_t" = "Externalizing T-score")) +
  theme_brand()

ggsave("output/WH_love.plot_after_matching.png",
       WH_love_plot, width = 8, height = 6)
```

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=14}
cowplot::plot_grid(WH_before_matching_plot, WH_after_matching_plot, WH_love_plot, nrow = 3)

WH_PSM <- cowplot::plot_grid(WH_before_matching_plot, WH_after_matching_plot, WH_love_plot, nrow = 3)

ggsave("output/WH_PSM.png",
       WH_PSM, width = 6, height = 9)
```

Merge and save matched dataset for White vs. Hispanic comparison.

```{r finalize-matched-wh}
WH_merged_data_sus <- susp_wide %>%
  filter(src_subject_id %in% WH_matched_data$src_subject_id) %>%
  left_join(WH_matched_data, by = "src_subject_id")

WH_merged_data_depression_sus <- depression_wide %>%
  filter(src_subject_id %in% WH_merged_data_sus$src_subject_id) %>%
  left_join(WH_merged_data_sus, by = "src_subject_id")

WH_matched_data_raw <- WH_merged_data_depression_sus %>%
  left_join(substance_wide, by = "src_subject_id") %>%
  left_join(grade_wide,     by = "src_subject_id") %>%
  select(-propensity_score, -distance, -weights, -subclass)

write.csv(WH_matched_data_raw,
          "data/WH_matched_data_raw.csv",
          row.names = FALSE)
cat("Saved: WH_matched_data_raw.csv\n")

# Pre-match dataset for White vs. Hispanic
WH_prematch_data_raw <- depression_wide %>%
  filter(src_subject_id %in% WH_baseline_ed_recode$src_subject_id) %>%
  left_join(WH_baseline_ed_recode, by = "src_subject_id") %>%
  left_join(substance_wide, by = "src_subject_id") %>%
  left_join(grade_wide,     by = "src_subject_id")

write.csv(WH_prematch_data_raw,
          "data/WH_prematch_data_raw.csv",
          row.names = FALSE)
cat("Saved:WH_prematch_data_raw.csv\n")

cat("White participants (control):",    sum(WH_prematch_data_raw$race_ethnicity_recode == 0), "\n")
cat("Hispanic participants (treated):", sum(WH_prematch_data_raw$race_ethnicity_recode == 1), "\n")

```


------------------------------------------------------------------------

# Descriptives Statistics

## Participant Characteristics -- Table 1

Summarizes participant characteristics for the overall sample.

::: callout-note
`render.missing = NULL` hides the "missing" row from the table.
:::

```{r table1-overall}
# ============================================================
# TABLE 1: FULL SAMPLE DESCRIPTIVE STATISTICS
# ============================================================

table_1 <- table1(
  ~ demo_brthdat_v2 + demo_sex_v2 + race_ethnicity + ledsch_seda_d_urban +
    kbi_p_c_school_setting + crpbi_caregiver1_y + parental_ed_bracket + income_bracket +
    marital_status_bracket + cbcl_scr_syn_external_t + cbcl_scr_dsm5_depress_t +
    cbcl_scr_syn_rulebreak_t + ed_det_susp_baseline_3,
  data = filtered_df_baseline_recoded,
  caption = "Sample Characteristics"
  #render.missing = NULL
)
table_1
```

We create separate Table 1s for White, Black, and Hispanic subgroups to show how sample characteristics differ across groups.

```{r table1-by-race}
# ============================================================
# TABLES 2-4: DESCRIPTIVE STATISTICS STRATIFIED BY RACE
# ============================================================

# --- White participants only ---
filtered_df_baseline_recoded_W <- filtered_df_baseline_recoded %>%
  filter(race_ethnicity == "White")

table_2 <- table1(
  ~ demo_brthdat_v2 + demo_sex_v2 + ledsch_seda_d_urban +
    kbi_p_c_school_setting + crpbi_caregiver1_y + parental_ed_bracket + income_bracket +
    marital_status_bracket + cbcl_scr_syn_external_t + cbcl_scr_dsm5_depress_t +
    cbcl_scr_syn_rulebreak_t + ed_det_susp_baseline_3,
  data = filtered_df_baseline_recoded_W,
  caption = "Characteristics: White Participants"
  #render.missing = NULL
)
table_2

# --- Black participants (including Multiracial Black) ---
filtered_df_baseline_recoded_B <- filtered_df_baseline_recoded %>%
  filter(as.numeric(race_ethnicity) %in% c(3, 6))

table_3 <- table1(
  ~ demo_brthdat_v2 + demo_sex_v2 + ledsch_seda_d_urban +
    kbi_p_c_school_setting + crpbi_caregiver1_y + parental_ed_bracket + income_bracket +
    marital_status_bracket + cbcl_scr_syn_external_t + cbcl_scr_dsm5_depress_t +
    cbcl_scr_syn_rulebreak_t  + ed_det_susp_baseline_3,
  data = filtered_df_baseline_recoded_B,
  caption = "Characteristics: Black Participants"
  #render.missing = NULL
)
table_3

# --- Hispanic participants (including Multiracial Hispanic) ---
filtered_df_baseline_recoded_H <- filtered_df_baseline_recoded %>%
  filter(as.numeric(race_ethnicity) %in% c(1, 5))

table_4 <- table1(
  ~ demo_brthdat_v2 + demo_sex_v2 + ledsch_seda_d_urban +
    kbi_p_c_school_setting + crpbi_caregiver1_y + parental_ed_bracket + income_bracket +
    marital_status_bracket + cbcl_scr_syn_external_t + cbcl_scr_dsm5_depress_t +
    cbcl_scr_syn_rulebreak_t +  ed_det_susp_baseline_3,
  data = filtered_df_baseline_recoded_H,
  caption = "Characteristics: Hispanic Participants"
  #render.missing = NULL
)
table_4
```

Convert the table objects to a flextable, then saves as .docx.

```{r save-table1}
# ============================================================
# SAVE TABLE'S TO A WORD DOCUMENT
# ============================================================

# table 1 (overall)
table_flex <- flextable(as.data.frame(table_1))

save_as_docx(
  `Table 1` = table_flex,
  path = "output/Table1.docx")
cat("Table 1 saved to: output/Table1.docx\n")

# table 2 (white)
table_flex2 <- flextable(as.data.frame(table_2))

save_as_docx(
  `Table 2` = table_flex2,
  path = "output/Table2.docx")
cat("Table 2 saved to: output/Table2.docx\n")

# table 3 (black)
table_flex3 <- flextable(as.data.frame(table_3))

save_as_docx(
  `Table 3` = table_flex3,
  path = "output/Table3.docx")
cat("Table 3 saved to: output/Table3.docx\n")

# table 4 (hispanic)
table_flex4 <- flextable(as.data.frame(table_4))

save_as_docx(
  `Table 4` = table_flex4,
  path = "output/Table4.docx")
cat("Table 4 saved to: output/Table4.docx\n")
```

What is the largest data collection site (reference group)

```{r site-descript-baseline, fig.width=15, fig.height=10}
filtered_df_baseline_recoded %>%
  ggplot(aes(x = site, fill = site)) + 
    geom_bar(color = "black") +
    geom_text(stat = "count", aes(label = after_stat(count)), 
              vjust = -0.5, size = 2) +
    scale_fill_brand() +
    labs(x = "Data Collection Site", fill = "Key") +
    theme_brand() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


------------------------------------------------------------------------

## Exclusionary Discipline Outcome (Year 3)

Proportions of ED at year 3 in the White-Black matched sample
```{r psm-wb-ed}
WB_matched_data_raw %>%
  mutate(race_ethnicity_recode = factor(race_ethnicity_recode, labels = c("White", "Black")),
         `DA at Year 3` = factor(ed_det_susp_baseline_3, labels =c("No", "Yes"))) %>%
  tbl_summary(include = `DA at Year 3`,
              by = race_ethnicity_recode,
              statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
  digits = all_continuous() ~ 2,
  missing_text = "(Missing)"
  ) %>%
  modify_spanning_header(all_stat_cols() ~ "**Race/ethnicity**") %>%
  add_p()
```

Proportions of ED at year 3 in the White-Hispanic matched sample
```{r psm-wh-ed}
WH_matched_data_raw  %>%
  mutate(race_ethnicity_recode = factor(race_ethnicity_recode, labels = c("White", "Hispanic")),
         `DA at Year 3` = factor(ed_det_susp_baseline_3, labels =c("No", "Yes"))) %>%
  tbl_summary(include = `DA at Year 3`,
              by = race_ethnicity_recode,
              statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
  digits = all_continuous() ~ 2,
  missing_text = "(Missing)"
  ) %>%
  modify_spanning_header(all_stat_cols() ~ "**Race/ethnicity**") %>%
  add_p()
```

------------------------------------------------------------------------

## Outcome Variables Across Follow-Up

Clean up `abcd_all` dataset for descriptive analyses
```{r recode-abcd-all}
abcd_all <- abcd_all %>%
    select(
    src_subject_id,
    site,                       # Data collection site
    eventname,                  # Data collection wave
    demo_brthdat_v2,            # Age (at visit, i.e., baseline)
    demo_sex_v2,                # Sex (1=Male, 2=Female)
    race_ethnicity,             # Recoded race/ethnicity
    demo_race_a_p___11,         # Black race
    demo_prnt_ed_v2_2yr_l,      # Parental educational attainment
    demo_prnt_income_v2,        # Household income bracket
    ledsch_seda_d_urban,        # Rural (1) vs. Non-rural (0)
    kbi_p_c_school_setting,     # School type (public/private/home)
    crpbi_caregiver1_y,         # Secondary caregiver support
    demo_prnt_marital_v2,       # Parental marital status
    cbcl_scr_syn_external_t,    # CBCL Externalizing T-score
    cbcl_scr_dsm5_depress_t,    # CBCL Depression T-score
    cbcl_scr_syn_rulebreak_t,   # CBCL Rule-Breaking T-score
    sag_grade_type,             # Grade in school
    kbi_p_c_det_susp_l,         # DA/ED
    tlfb_cal_scr_tot_su_ud,     # Substance use number of days 
   #cbcl_scr_syn_rulebreak_sum, # CBCL Rule-Breaking Sum 
    cbcl_q28_p,                 # CBCL item 28
    cbcl_q23_p                  # CBCL item 23
  ) %>%
  mutate(race_ethnicity = case_when(
    race_ethnicity == 1                                  ~ 1, # hispanic
    race_ethnicity == 2                                  ~ 2, # white
    race_ethnicity == 3                                  ~ 3, # black
    race_ethnicity == 7                                  ~ 4, # asian/pacific islander
    race_ethnicity == 11                                 ~ 5, # multiracial + hispanic
    race_ethnicity == 12 & demo_race_a_p___11 == 1       ~ 6, # multiracial + black
    race_ethnicity == 12 & (demo_race_a_p___11 != 1
                            | is.na(demo_race_a_p___11)) ~ 7, # other
    race_ethnicity == 13                                 ~ 7, # other
    TRUE                                                 ~ race_ethnicity
  )) %>%
  mutate(
    income_bracket = case_when(
      demo_prnt_income_v2 %in% 1:2   ~ 1,        # < $50,000
      demo_prnt_income_v2 == 3       ~ 2,        # $50,000 - $74,999
      demo_prnt_income_v2 == 4       ~ 3,        # $75,000 - $99,999
      demo_prnt_income_v2 == 5       ~ 4,        # $100,000 - $199,999
      demo_prnt_income_v2 == 6       ~ 5,        # >= $200,000
      demo_prnt_income_v2 == 777     ~ 777,      # Refused
      demo_prnt_income_v2 == 999     ~ 999,      # Don't know
      TRUE                           ~ NA_real_
    )
  ) %>%
  mutate(
    parental_ed_bracket = case_when(
      demo_prnt_ed_v2_2yr_l == 777                 ~ NA_real_, # Refused
      demo_prnt_ed_v2_2yr_l == 999                 ~ NA_real_, # Don't know
      TRUE                                         ~ demo_prnt_ed_v2_2yr_l
    )
  ) %>%
  mutate(
    marital_status_bracket = case_when(
      demo_prnt_marital_v2 %in% c(1, 6)    ~ 1,        # Married or living with partner
      demo_prnt_marital_v2 %in% c(2, 3, 4) ~ 2,        # Widowed / Divorced / Separated
      demo_prnt_marital_v2 == 5            ~ 3,        # Never married
      demo_prnt_marital_v2 == 777          ~ NA_real_, # Refused
      TRUE                                 ~ NA_real_
    )
  ) %>% 
    mutate(
    kbi_p_c_det_susp_l = case_when(
      kbi_p_c_det_susp_l == 999   ~ NA_real_,  # not sure
      kbi_p_c_det_susp_l == 777   ~ NA_real_,  # declined to answer
      TRUE                        ~ kbi_p_c_det_susp_l
    )
  )

# Convert urban, marital status and caregiver to factor (categorical) type
abcd_all$ledsch_seda_d_urban <- as.factor(abcd_all$ledsch_seda_d_urban)
abcd_all$crpbi_caregiver1_y  <- as.factor(abcd_all$crpbi_caregiver1_y)
filtered_df_baseline$marital_status_bracket <- as.factor(filtered_df_baseline$marital_status_bracket)

# Academic grades: convert letter grade codes to GPA-equivalent numeric values
# Using the lower boundary for each grade range
abcd_all <- abcd_all%>%
  mutate(
    grade_gpa = case_when(
      sag_grade_type == 1  ~ 4.0,  # A+
      sag_grade_type == 2  ~ 3.7,  # A
      sag_grade_type == 3  ~ 3.5,  # A-
      sag_grade_type == 4  ~ 3.3,  # B+
      sag_grade_type == 5  ~ 3.0,  # B
      sag_grade_type == 6  ~ 2.7,  # B-
      sag_grade_type == 7  ~ 2.3,  # C+
      sag_grade_type == 8  ~ 2.0,  # C
      sag_grade_type == 9  ~ 1.7,  # C-
      sag_grade_type == 10 ~ 1.3,  # D+
      sag_grade_type == 11 ~ 1.0,  # D
      sag_grade_type == 12 ~ 0.0,  # F
      TRUE                 ~ NA_real_
    )
  )

# Create a binary variable indicating any substance use that year
abcd_all <- abcd_all %>%
  mutate(any_substance = ifelse(tlfb_cal_scr_tot_su_ud > 0, 1, 0))
```

```{r tbl-outcome-fu}
abcd_all %>%
  mutate(
    `Depression (CBCL) t-score` = cbcl_scr_dsm5_depress_t, 
    `Any substance use`             = any_substance,
    `Days of substance use`     = tlfb_cal_scr_tot_su_ud,
    `Academic grade (GPA)`      = grade_gpa,
    Wave = factor(eventname, labels = c("Baseline", "1 Year", "2 Year", "3 Year", "4 Year", "5 Year", "6 Year", "7 Year"))
  ) %>%
  tbl_summary(
    include = c(`Depression (CBCL) t-score`, `Any substance use`, `Days of substance use`, `Academic grade (GPA)`),
    by      = Wave,
    type    = all_continuous() ~ "continuous2", 
    statistic = list(
      all_continuous()  ~ c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}"),
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits       = all_continuous() ~ 2,
    missing_text = "(Missing)"
  ) %>%
  modify_spanning_header(
    c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5", "stat_6", "stat_7") ~ "**Data Collection Wave**"
  )

# save as doc 
abcd_all %>%
  mutate(
    `Depression (CBCL) t-score` = cbcl_scr_dsm5_depress_t, 
    `Any substance use`             = any_substance,
    `Days of substance use`     = tlfb_cal_scr_tot_su_ud,
    `Academic grade (GPA)`      = grade_gpa,
    Wave = factor(eventname, labels = c("Baseline", "1 Year", "2 Year", "3 Year", "4 Year", "5 Year", "6 Year", "7 Year"))
  ) %>%
  tbl_summary(
    include = c(`Depression (CBCL) t-score`, `Any substance use`, `Days of substance use`, `Academic grade (GPA)`),
    by      = Wave,
    type    = all_continuous() ~ "continuous2", 
    statistic = list(
      all_continuous()  ~ c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}"),
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits       = all_continuous() ~ 2,
    missing_text = "(Missing)"
  ) %>%
  modify_spanning_header(
    c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5", "stat_6", "stat_7") ~ "**Data Collection Wave**"
  ) %>%
  as_gt() %>%
  gt::gtsave(filename = "output/outcome-across-fu.docx")
```

```{r ed-across-fu}
abcd_all %>%
  mutate(
    `Disciplinary action` = kbi_p_c_det_susp_l, 
    Wave = factor(eventname, labels = c("Baseline", "1 Year", "2 Year", "3 Year", "4 Year", "5 Year", "6 Year", "7 Year"))
  ) %>%
  tbl_summary(
    include = c(`Disciplinary action`),
    by      = Wave,
    type    = all_continuous() ~ "continuous2", 
    statistic = list(
      all_continuous()  ~ c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}"),
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits       = all_continuous() ~ 2,
    missing_text = "(Missing)"
  ) %>%
  modify_spanning_header(
    c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5", "stat_6", "stat_7") ~ "**Data Collection Wave**"
  )
```


# Linear Regression Analyses

## Section A: Pre-match Regressions

Prematch analyses includes the whole cohort.

### A1: Depression -- Pre-match (WB)

OUTCOME: Depression T-score at 4-year follow-up

EXPOSURE: Ever suspended by year 3 (binary: 0/1)

SAMPLE: White + Black, UNMATCHED (full pre-match sample)

COVARIATE: Baseline depression T-score used as a control so we're modeling [change]{.underline} in depression, not level.

Four models:

1.  Everyone pooled (all White + Black participants)

2.  White participants only

3.  Black participants only

4.  Pooled model with race (is_black | is_his)

EXPLORATORY: Null model with race x suspension interaction (no other covariates)

```{r prematch-wb-depression-binary}

# Ensure is_black indicator is set (1=Black/Multiracial Black, 0=White)
WB_prematch_data_raw <- WB_prematch_data_raw %>%
  mutate(is_black = ifelse(race_ethnicity %in% c(3, 6), 1, 0))

# Select only the variables needed for this set of models
regression_subset <- WB_prematch_data_raw[, c(
  "cbcl_scr_syn_external_t",             # CBCL externalising-score (covariate)
  "cbcl_scr_dsm5_depress_t_ses-00A",     # Baseline depression (control)
  "cbcl_scr_dsm5_depress_t_ses-04A",     # Depression at year 4 (outcome)
  "ed_det_susp_baseline_3",              # Ever suspended (key exposure)
  "ed_det_susp_count",                   # Suspension count (used in ordinal models below)
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_black",
  "site"
)]

regression_subset <- regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Model 1: Full pooled sample (White + Black combined), controlling for covariates
# lm() = linear regression; missing rows are automatically dropped (listwise deletion)
prematch_depress_model_WB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_depress_model_WB)

# Model 2: White participants only (is_black == 0)
regression_subset_white <- regression_subset %>% filter(is_black == 0)

prematch_depress_model_onlyWB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_white
)
summary(prematch_depress_model_onlyWB)

# Model 3: Black participants only (is_black == 1)
regression_subset_black <- regression_subset %>% filter(is_black == 1)

prematch_depress_model_onlyB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_black
)
summary(prematch_depress_model_onlyB)

# Model 4: Pooled model with race control
prematch_depress_race_control_WB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_depress_race_control_WB)

# Null model: Tests whether suspension effect DIFFERS by race (interaction term)
# is_black * ed_det_susp_baseline_3 = the interaction (does race moderate the effect?)
prematch_depress_model_WB_null <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + is_black * ed_det_susp_baseline_3,
  data = regression_subset
)
summary(prematch_depress_model_WB_null)
```

### A1b: Depression -- Ordinal Suspension Count (WB Pre-match)

Same models as above, but using ed_det_susp_count (0, 1, or 2) instead of the binary ever-suspended variable. This tests whether MORE suspensions = worse outcomes.

```{r prematch-wb-depression-ordinal}

describe(regression_subset$ed_det_susp_count)

# Pooled model
prematch_depress_model_count <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_count + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_depress_model_count)

# Pooled with race control
prematch_depress_race_count <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_black + ed_det_susp_count + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_depress_race_count)

# White only
regression_subset_white <- regression_subset %>% filter(is_black == 0)
prematch_depress_model_onlyW_count <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_count + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_white
)
summary(prematch_depress_model_onlyW_count)

# Black only
regression_subset_black <- regression_subset %>% filter(is_black == 1)
prematch_depress_model_onlyB_count <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_count + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_black
)
summary(prematch_depress_model_onlyB_count)

# Null model: Tests whether suspension effect DIFFERS by race (interaction term)
prematch_depress_model_WB_count_null <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_black + ed_det_susp_count + is_black * ed_det_susp_count,
  data = regression_subset
)
summary(prematch_depress_model_WB_count_null)
```

### A2: Depression -- Pre-match (WH)

```{r prematch-wh-depression}
# ============================================================
# Same depression models, now for the White vs. Hispanic sample.
# is_his = 1 for Hispanic/Multiracial Hispanic, 0 for White.
# ============================================================

WH_prematch_data_raw <- WH_prematch_data_raw %>%
  mutate(is_his = ifelse(race_ethnicity %in% c(1, 5), 1, 0))

regression_subset <- WH_prematch_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "cbcl_scr_dsm5_depress_t_ses-00A",
  "cbcl_scr_dsm5_depress_t_ses-04A", # Depression at year 4 (outcome)
  "ed_det_susp_baseline_3",
  "ed_det_susp_count",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_his",
  "site"
)]

regression_subset <- regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled model (all White + Hispanic)
prematch_depress_model_WH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_depress_model_WH)

# White only (is_his == 0)
regression_subset_white <- regression_subset %>% filter(is_his == 0)
prematch_depress_model_onlyWH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_white
)
summary(prematch_depress_model_onlyWH)

# Hispanic only (is_his == 1)
regression_subset_his <- regression_subset %>% filter(is_his == 1)
prematch_depress_model_onlyH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_his
)
summary(prematch_depress_model_onlyH)

# Pooled with race control
prematch_depress_model_race_control_WH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_depress_model_race_control_WH)

# Null model: Tests whether suspension effect DIFFERS by race (interaction term)
prematch_depress_model_WH_null <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + is_his * ed_det_susp_baseline_3,
  data = regression_subset_his
)
summary(prematch_depress_model_WH_null)
```

### A3: Substance Use -- Pre-match (WB)

 OUTCOME: Total substance use disorder days at 4-year follow-up
 
 EXPOSURE: Ever suspended by year 3 (binary)
 
 COVARIATE: Total substance use disorder at baseline used as a control

```{r prematch-wb-substance}
regression_subset <- WB_prematch_data_raw[, c(
  "cbcl_scr_syn_external_t",         # Externalising t-score (covariate)
  "tlfb_cal_scr_tot_su_ud_ses-00A",  # Baseline substance use (control)
  "tlfb_cal_scr_tot_su_ud_ses-04A",  # Substance use at year 4 (outcome)
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_black",
  "site"
)]

regression_subset <- regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled model
prematch_SU_model_WB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_SU_model_WB)

# White only
regression_subset_white <- regression_subset %>% filter(is_black == 0)
prematch_SU_model_onlyWB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_white
)
summary(prematch_SU_model_onlyWB)

# Black only
regression_subset_black <- regression_subset %>% filter(is_black == 1)
prematch_SU_model_onlyB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_black
)
summary(prematch_SU_model_onlyB)

# Pooled with race control
prematch_SU_model_race_control_WB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_SU_model_race_control_WB)

# Null model: Tests whether suspension effect DIFFERS by race (interaction term)
prematch_SU_model_WB_null <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + is_black * ed_det_susp_baseline_3,
  data = regression_subset
)
summary(prematch_SU_model_WB_null)
```

### A4: Substance Use -- Pre-match (WH)

```{r prematch-wh-substance}
regression_subset <- WH_prematch_data_raw[, c(
  "cbcl_scr_syn_external_t",         # Externalising-score (covariate)
  "tlfb_cal_scr_tot_su_ud_ses-00A",  # Baseline substance use (control)
  "tlfb_cal_scr_tot_su_ud_ses-04A",  # Substance use at year 4 (outcome)
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_his",
  "site"
)]

regression_subset <- regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled model
prematch_SU_model_WH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_SU_model_WH)

# White only
regression_subset_white <- regression_subset %>% filter(is_his == 0)
prematch_SU_model_onlyWH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_white
)
summary(prematch_SU_model_onlyWH)

# Black only
regression_subset_his <- regression_subset %>% filter(is_his == 1)
prematch_SU_model_onlyH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_his
)
summary(prematch_SU_model_onlyH)

# Pooled with race control
prematch_SU_model_race_control_WH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_SU_model_race_control_WH)

# Null model: Tests whether suspension effect DIFFERS by race (interaction term)
prematch_SU_model_WH_null <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + is_his * ed_det_susp_baseline_3,
  data = regression_subset
)
summary(prematch_SU_model_WH_null)
```

### A5: Academic Grades -- Pre-match (WB)

OUTCOME: GPA at 4-year follow-up

EXPOSURE: Ever suspended by year 3 (binary)

COVARIATE: GPA at year 3 used as a control (so we model change in GPA, not absolute GPA)

```{r prematch-wb-grade}
regression_subset <- WB_prematch_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "grade_gpa_ses-03A",  # Year 3 GPA (control)
  "grade_gpa_ses-04A",  # Year 4 GPA (outcome)
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_black",
  "ed_det_susp_count",
  "site"
)]

regression_subset <- regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled
prematch_grade_model_WB <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_grade_model_WB)

# White only
regression_subset_white <- regression_subset %>% filter(is_black == 0)
prematch_grade_model_onlyWB <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_white
)
summary(prematch_grade_model_onlyWB)

# Black only
regression_subset_black <- regression_subset %>% filter(is_black == 1)
prematch_grade_model_onlyB <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_black
)
summary(prematch_grade_model_onlyB)

# Pooled with race control
prematch_grade_model_race_control_WB <- lm(
  `grade_gpa_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_grade_model_race_control_WB)

# Null model: Tests whether suspension effect DIFFERS by race (interaction term)
prematch_grade_model_null_WB <- lm(
  `grade_gpa_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + is_black * ed_det_susp_baseline_3,
  data = regression_subset
)
summary(prematch_grade_model_null_WB)
```

### A6: Academic Grades -- Pre-match (WH)

```{r prematch-wh-grade}
# Same grade models for the White vs. Hispanic (pre-match) sample

regression_subset <- WH_prematch_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "grade_gpa_ses-03A",  # Year 3 GPA (control)
  "grade_gpa_ses-04A",  # Year 4 GPA (outcome)
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_his",
  "ed_det_susp_count",
  "site"
)]

regression_subset <- regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled
prematch_grade_model_WH <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_grade_model_WH)

# White only (is_his == 0)
regression_subset_white <- regression_subset %>% filter(is_his == 0)
prematch_grade_model_onlyWH <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_white
)
summary(prematch_grade_model_onlyWH)

# Hispanic only (is_his == 1)
regression_subset_his <- regression_subset %>% filter(is_his == 1)
prematch_grade_model_onlyH <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset_his
)
summary(prematch_grade_model_onlyH)

# Pooled with race control
prematch_grade_model_race_control_WH <- lm(
  `grade_gpa_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = regression_subset
)
summary(prematch_grade_model_race_control_WH)

# Null model: Tests whether suspension effect DIFFERS by race (interaction term)
prematch_grade_model_null_WH <- lm(
  `grade_gpa_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + is_his * ed_det_susp_baseline_3,
  data = regression_subset
)
summary(prematch_grade_model_null_WH)
```

## Section B: Matched Data Regressions

### B1: Depression -- Matched (WB)

Same depression models as Section A, but now using the **PROPENSITY-SCORE-MATCHED** sample. The matched sample has been balanced on confounders, so any remaining differences should \# better reflect the causal effect of suspension.

```{r matched-wb-depression}

WB_matched_data_raw <- WB_matched_data_raw %>%
  mutate(is_black = ifelse(race_ethnicity %in% c(3, 6), 1, 0))

matched_regression_subset <- WB_matched_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "cbcl_scr_dsm5_depress_t_ses-00A",
  "cbcl_scr_dsm5_depress_t_ses-04A",
  "ed_det_susp_baseline_3",
  "ed_det_susp_count",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_black",
  "site"
)]

matched_regression_subset <- matched_regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled matched
matched_depress_model_WB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_depress_model_WB)

# White only (matched)
matched_regression_subset_white <- matched_regression_subset %>% filter(is_black == 0)
matched_depress_model_onlyWB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_white
)
summary(matched_depress_model_onlyWB)

# Black only (matched)
matched_regression_subset_black <- matched_regression_subset %>% filter(is_black == 1)
matched_depress_model_onlyB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_black
)
summary(matched_depress_model_onlyB)

# Pooled matched with race control
matched_depress_model_race_control_WB <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_depress_model_race_control_WB)

# Null model: Tests whether suspension effect DIFFERS by race (matched sample)
matched_depress_model_WB_null <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + is_black * ed_det_susp_baseline_3,
  data = matched_regression_subset
)
summary(matched_depress_model_WB_null)
```

### B2: Depression -- Matched (WH)

Same depression models for the White vs. Hispanic matched sample

```{r matched-wh-depression}

WH_matched_data_raw <- WH_matched_data_raw %>%
  mutate(is_his = ifelse(race_ethnicity %in% c(1, 5), 1, 0))

matched_regression_subset <- WH_matched_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "cbcl_scr_dsm5_depress_t_ses-00A",
  "cbcl_scr_dsm5_depress_t_ses-04A",
  "ed_det_susp_baseline_3",
  "ed_det_susp_count",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_his",
  "site"
)]

matched_regression_subset <- matched_regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled
matched_depress_model_WH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_depress_model_WH)

# White only (is_his == 0)
matched_regression_subset_white <- matched_regression_subset %>% filter(is_his == 0)
matched_depress_model_onlyWH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_white
)
summary(matched_depress_model_onlyWH)

# Hispanic only (is_his == 1)
matched_regression_subset_his <- matched_regression_subset %>% filter(is_his == 1)
matched_depress_model_onlyH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_his
)
summary(matched_depress_model_onlyH)

# Pooled with race control
matched_depress_model_race_control_WH <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + `cbcl_scr_dsm5_depress_t_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_depress_model_race_control_WH)

# Null model: Tests whether suspension effect DIFFERS by race (matched sample)
matched_depress_model_WH_null <- lm(
  `cbcl_scr_dsm5_depress_t_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + is_his * ed_det_susp_baseline_3,
  data = matched_regression_subset_his
)
summary(matched_depress_model_WH_null)
```

### B3: Substance Use -- Matched (WB)

Substance use regression on the White vs. Black MATCHED sample

```{r matched-wb-substance}

matched_regression_subset <- WB_matched_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "tlfb_cal_scr_tot_su_ud_ses-00A",
  "tlfb_cal_scr_tot_su_ud_ses-04A",
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_black",
  "site"
)]

matched_regression_subset <- matched_regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled model
matched_SU_model_WB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_SU_model_WB)


# White only
matched_regression_subset_white <- matched_regression_subset %>% filter(is_black == 0)
matched_SU_model_onlyWB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_white
)
summary(matched_SU_model_onlyWB)

# Black only
matched_regression_subset_black <- matched_regression_subset %>% filter(is_black == 1)
matched_SU_model_onlyB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_black
)
summary(matched_SU_model_onlyB)

# Pooled with race control 
matched_SU_model_race_control_WB <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_SU_model_race_control_WB)

# Null model: Tests whether suspension effect DIFFERS by race (matched sample)
matched_SU_model_WB_null <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + is_black * ed_det_susp_baseline_3,
  data = matched_regression_subset
)
summary(matched_depress_model_WB_null)
```

### B4: Substance Use -- Matched (WH)

Substance use regression on the White vs. Hispanic MATCHED sample

```{r matched-wh-substance}

matched_regression_subset <- WH_matched_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "tlfb_cal_scr_tot_su_ud_ses-00A",
  "tlfb_cal_scr_tot_su_ud_ses-04A",
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_his",
  "site"
)]

matched_regression_subset <- matched_regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled model
matched_SU_model_WH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_SU_model_WH)


# White only
matched_regression_subset_white <- matched_regression_subset %>% filter(is_his == 0)
matched_SU_model_onlyWH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_white
)
summary(matched_SU_model_onlyWH)

# Hispanic only
matched_regression_subset_his <- matched_regression_subset %>% filter(is_his == 1)
matched_SU_model_onlyH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_his
)
summary(matched_SU_model_onlyH)

# Pooled with race control 
matched_SU_model_race_control_WH <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + `tlfb_cal_scr_tot_su_ud_ses-00A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_SU_model_race_control_WH)

# Null model: Tests whether suspension effect DIFFERS by race (matched sample)
matched_SU_model_WH_null <- lm(
  `tlfb_cal_scr_tot_su_ud_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + is_his * ed_det_susp_baseline_3,
  data = matched_regression_subset
)
summary(matched_depress_model_WH_null)
```

### B5: Academic Grades -- Matched (WB)

Grade regression on the White vs. Black MATCHED sample

```{r matched-wb-grade}

matched_regression_subset <- WB_matched_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "grade_gpa_ses-03A",
  "grade_gpa_ses-04A",
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_black",
  "ed_det_susp_count",
  "site"
)]

matched_regression_subset <- matched_regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

# Pooled
matched_grade_model_WB <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_grade_model_WB)

# White only
matched_regression_subset_white <- matched_regression_subset %>% filter(is_black == 0)
matched_grade_model_onlyWB <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_white
)
summary(matched_grade_model_WB)

# Black only
matched_regression_subset_black <- matched_regression_subset %>% filter(is_black == 1)
matched_grade_model_onlyB <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_black
)
summary(matched_grade_model_onlyB)

# Pooled with race control
matched_grade_model_race_control_WB <- lm(
  `grade_gpa_ses-04A` ~
    is_black + ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_grade_model_race_control_WB)

# Null model: Tests whether suspension effect DIFFERS by race (matched sample)
matched_grade_model_null <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + is_black + is_black * ed_det_susp_baseline_3,
  data = matched_regression_subset
)
summary(matched_grade_model_null)
```

### B6: Academic Grades -- Matched (WH)

Grade regression on the White vs. Hispanic MATCHED sample

```{r matched-wh-grade}

matched_regression_subset <- WH_matched_data_raw[, c(
  "cbcl_scr_syn_external_t",
  "grade_gpa_ses-03A",
  "grade_gpa_ses-04A",
  "ed_det_susp_baseline_3",
  "demo_sex_v2",
  "crpbi_caregiver1_y",
  "income_bracket",
  "parental_ed_bracket",
  "marital_status_bracket",
  "race_ethnicity",
  "race_ethnicity_recode",
  "is_his",
  "ed_det_susp_count",
  "site"
)]

matched_regression_subset <- matched_regression_subset %>%
  mutate(income_bracket = as.factor(income_bracket),
         marital_status_bracket = as.factor(marital_status_bracket),
         parental_ed_bracket = as.factor(parental_ed_bracket))

sapply(matched_regression_subset, class) # check type to confirm factor/numeric assignment

# Pooled
matched_grade_model_WH <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset
)
summary(matched_grade_model_WH)

# White only (is_his == 0)
matched_regression_subset_white <- matched_regression_subset %>% filter(is_his == 0)
matched_grade_model_onlyWH <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_white
)
summary(matched_grade_model_onlyWH)

# Hispanic only (is_his == 1)
matched_regression_subset_his <- matched_regression_subset %>% filter(is_his == 1)
matched_grade_model_onlyH <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_his
)
summary(matched_grade_model_onlyH)

# Pooled with race control
matched_grade_model_race_control_WH <- lm(
  `grade_gpa_ses-04A` ~
    is_his + ed_det_susp_baseline_3 + `grade_gpa_ses-03A` + site +
    demo_sex_v2 + crpbi_caregiver1_y + income_bracket +
    parental_ed_bracket + marital_status_bracket + cbcl_scr_syn_external_t,
  data = matched_regression_subset_his
)
summary(matched_grade_model_race_control_WH)

# Null model: Tests whether suspension effect DIFFERS by race (matched sample)
matched_grade_model_null <- lm(
  `grade_gpa_ses-04A` ~
    ed_det_susp_baseline_3 + is_his + is_his * ed_det_susp_baseline_3,
  data = matched_regression_subset_his
)
summary(matched_grade_model_null)
```

------------------------------------------------------------------------

## Section C: Export Regression Table to Word

```{r helper-get-all-variables}
# ============================================================
# HELPER FUNCTION: get_all_variables()
#
# Collects all unique variable names (predictors) across a list
# of models. Used to build a combined results table where each
# row is a predictor and each column is a different model.
# ============================================================
get_all_variables <- function(model_list) {
  all_vars <- unique(unlist(lapply(model_list, function(m) rownames(coef(summary(m))))))
  return(all_vars)
}

format_coef <- function(estimate, p_value) {
  formatted <- sprintf("%.3f", estimate)
  if      (p_value < 0.001) formatted <- paste0(formatted, "***")
  else if (p_value < 0.01)  formatted <- paste0(formatted, "**")
  else if (p_value < 0.05)  formatted <- paste0(formatted, "*")
  return(formatted)
}
```

### C1: Depression -- Matched (WB)

```{r build-depress-regression-table-wb}
# ── Define model list ─────────────────────────────────────────────────────────
WB_model_depress_full_sample <- list(
  "All Participants" = prematch_depress_model_WB,
  "White Only"       = prematch_depress_model_onlyWB,
  "Black Only"       = prematch_depress_model_onlyB,
  "Race-Controlled"  = prematch_depress_race_control_WB
)

WB_model_depress_full_matched <- list(
  "All Participants" = matched_depress_model_WB,
  "White Only"       = matched_depress_model_onlyWB,
  "Black Only"       = matched_depress_model_onlyB,
  "Race-Controlled"  = matched_depress_model_race_control_WB
)

# ── Shared reference levels ───────────────────────────────────────────────────

reference_levels <- c(
  "(Intercept)"                     = "N/A",
  "is_black"                        = "Non-Black (0)",
  "ed_det_susp_baseline_31"         = "No Suspension (0)",
  "cbcl_scr_dsm5_depress_t_ses-00A" = "Continuous Score",
  "demo_sex_v22"                    = "Male (1)",
  "crpbi_caregiver1_y1"             = "Continuous Score",
  "income_bracket2"                 = "<$50K (Reference)",
  "income_bracket3"                 = "<$50K (Reference)",
  "income_bracket4"                 = "<$50K (Reference)",
  "income_bracket5"                 = "<$50K (Reference)",
  "income_bracket777"               = "<$50K (Reference)",
  "income_bracket999"               = "<$50K (Reference)",
  "parental_ed_bracket2"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket3"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket4"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket5"            = "Up to HS no diploma (Reference)",
  "marital_status_bracket2"         = "Married/Living Together (Reference)",
  "marital_status_bracket3"         = "Married/Living Together (Reference)",
  "siteFlorida International University"        = "Children's Hospital Los Angeles",
  "siteLaureate Institute for Brain Research"   = "Children's Hospital Los Angeles",
  "siteMedical University of South Carolina"    = "Children's Hospital Los Angeles",
  "siteOregon Health & Science University"      = "Children's Hospital Los Angeles",
  "siteSRI International"                       = "Children's Hospital Los Angeles",
  "siteUniversity of California, Los Angeles"   = "Children's Hospital Los Angeles",
  "siteUniversity of California, San Diego"     = "Children's Hospital Los Angeles",
  "siteUniversity of Colorado Boulder"          = "Children's Hospital Los Angeles",
  "siteUniversity of Florida"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Maryland, Baltimore"       = "Children's Hospital Los Angeles",
  "siteUniversity of Michigan"                  = "Children's Hospital Los Angeles",
  "siteUniversity of Minnesota"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Pittsburgh Medical Center" = "Children's Hospital Los Angeles",
  "siteUniversity of Rochester"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Utah"                      = "Children's Hospital Los Angeles",
  "siteUniversity of Vermont"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Wisconsin, Milwaukee"      = "Children's Hospital Los Angeles",
  "siteVirginia Commonwealth University"        = "Children's Hospital Los Angeles",
  "siteWashington University in St. Louis"      = "Children's Hospital Los Angeles",
  "siteYale University"                         = "Children's Hospital Los Angeles"
)

# ── Core builder function ─────────────────────────────────────────────────────

build_coef_table <- function(model_list, title_label) {
  
  all_variable_names <- get_all_variables(model_list)
  
  # Auto-fill any missing reference levels
  ref_levels <- reference_levels  # copy shared reference levels
  for (var in all_variable_names) {
    if (!(var %in% names(ref_levels))) {
      if      (grepl("income_bracket",         var)) ref_levels[var] <- "<$50K (Reference)"
      else if (grepl("parental_ed_bracket",  var)) ref_levels[var] <- "Up to HS no diploma (Reference)"
      else if (grepl("marital_status_bracket", var)) ref_levels[var] <- "Married/Living Together (Reference)"
      else                                            ref_levels[var] <- "Reference Category"
    }
  }
  
  # Base data frame
  all_coefs <- data.frame(
    Variable  = all_variable_names,
    Reference = ref_levels[all_variable_names],
    stringsAsFactors = FALSE
  )
  
  # Fill in coefficients from each model
  for (model_name in names(model_list)) {
    model         <- model_list[[model_name]]
    model_summary <- summary(model)
    coefs         <- coef(model_summary)
    
    all_coefs[[model_name]] <- NA
    
    for (i in 1:nrow(all_coefs)) {
      var_name <- all_coefs$Variable[i]
      if (var_name %in% rownames(coefs)) {
        estimate <- coefs[var_name, "Estimate"]
        p_value  <- coefs[var_name, "Pr(>|t|)"]
        all_coefs[i, model_name] <- format_coef(estimate, p_value)
      }
    }
  }
  
  # Build flextable
  n_models <- length(model_list)
  
  ft <- flextable(all_coefs) %>%
    add_header_row(
      values    = c("", "", title_label),
      colwidths = c(1, 1, n_models)
    ) %>%
    theme_booktabs() %>%
    align(j = 1:2,              align = "left") %>%
    align(j = 3:(2 + n_models), align = "center") %>%
    width(j = 1,              width = 3.5) %>%
    width(j = 2,              width = 2.0) %>%
    width(j = 3:(2 + n_models), width = 1.5) %>%
    bold(i = 1, bold = TRUE) %>%
    add_footer_lines(values = c(
      "* p < 0.05, ** p < 0.01, *** p < 0.001",
      "Values show regression coefficients with significance stars",
      "Outcome: Depression symptoms (CBCL DSM-5 T-Score) at 3-year follow-up"
    ))
  
  return(ft)
}

# ── Generate tables ───────────────────────────────────────────────────────────

WB_depress_sample  <- build_coef_table(
  WB_model_depress_full_sample,
  "Depression (CBCL DSM-5 Depression T-Score) — Pre-Matched Sample"
)

WB_depress_matched <- build_coef_table(
  WB_model_depress_full_matched,
  "Depression (CBCL DSM-5 Depression T-Score) — Matched Sample"
)

WB_depress_sample
WB_depress_matched
```

### C2: Depression -- Matched (WH)

```{r build-depress-regression-table-wh}
# ── Define model list ─────────────────────────────────────────────────────────
WH_model_depress_full_sample <- list(
  "All Participants" = prematch_depress_model_WH,
  "White Only"       = prematch_depress_model_onlyWH,
  "Hispanic Only"    = prematch_depress_model_onlyH,
  "Race-Controlled"  = prematch_depress_model_race_control_WH
)

WH_model_depress_full_matched <- list(
  "All Participants" = matched_depress_model_WH,
  "White Only"       = matched_depress_model_onlyWH,
  "Hispanic Only"    = matched_depress_model_onlyH,
  "Race-Controlled"  = matched_depress_model_race_control_WH
)

# ── Shared reference levels ───────────────────────────────────────────────────

reference_levels <- c(
  "(Intercept)"                     = "N/A",
  "is_his"                          = "Non-Hispanic (0)",
  "ed_det_susp_baseline_31"         = "No Suspension (0)",
  "cbcl_scr_dsm5_depress_t_ses-00A" = "Continuous Score",
  "demo_sex_v22"                    = "Male (1)",
  "crpbi_caregiver1_y1"             = "Continuous Score",
  "income_bracket2"                 = "<$50K (Reference)",
  "income_bracket3"                 = "<$50K (Reference)",
  "income_bracket4"                 = "<$50K (Reference)",
  "income_bracket5"                 = "<$50K (Reference)",
  "income_bracket777"               = "<$50K (Reference)",
  "income_bracket999"               = "<$50K (Reference)",
  "parental_ed_bracket2"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket3"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket4"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket5"            = "Up to HS no diploma (Reference)",
  "marital_status_bracket2"         = "Married/Living Together (Reference)",
  "marital_status_bracket3"         = "Married/Living Together (Reference)",
  "siteFlorida International University"        = "Children's Hospital Los Angeles",
  "siteLaureate Institute for Brain Research"   = "Children's Hospital Los Angeles",
  "siteMedical University of South Carolina"    = "Children's Hospital Los Angeles",
  "siteOregon Health & Science University"      = "Children's Hospital Los Angeles",
  "siteSRI International"                       = "Children's Hospital Los Angeles",
  "siteUniversity of California, Los Angeles"   = "Children's Hospital Los Angeles",
  "siteUniversity of California, San Diego"     = "Children's Hospital Los Angeles",
  "siteUniversity of Colorado Boulder"          = "Children's Hospital Los Angeles",
  "siteUniversity of Florida"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Maryland, Baltimore"       = "Children's Hospital Los Angeles",
  "siteUniversity of Michigan"                  = "Children's Hospital Los Angeles",
  "siteUniversity of Minnesota"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Pittsburgh Medical Center" = "Children's Hospital Los Angeles",
  "siteUniversity of Rochester"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Utah"                      = "Children's Hospital Los Angeles",
  "siteUniversity of Vermont"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Wisconsin, Milwaukee"      = "Children's Hospital Los Angeles",
  "siteVirginia Commonwealth University"        = "Children's Hospital Los Angeles",
  "siteWashington University in St. Louis"      = "Children's Hospital Los Angeles",
  "siteYale University"                         = "Children's Hospital Los Angeles"
)

# ── Core builder function ─────────────────────────────────────────────────────

build_coef_table <- function(model_list, title_label) {
  
  all_variable_names <- get_all_variables(model_list)
  
  # Auto-fill any missing reference levels
  ref_levels <- reference_levels  # copy shared reference levels
  for (var in all_variable_names) {
    if (!(var %in% names(ref_levels))) {
      if      (grepl("income_bracket",         var)) ref_levels[var] <- "<$50K (Reference)"
      else if (grepl("parental_ed_bracket",  var)) ref_levels[var] <- "Up to HS no diploma (Reference)"
      else if (grepl("marital_status_bracket", var)) ref_levels[var] <- "Married/Living Together (Reference)"
      else                                            ref_levels[var] <- "Reference Category"
    }
  }
  
  # Base data frame
  all_coefs <- data.frame(
    Variable  = all_variable_names,
    Reference = ref_levels[all_variable_names],
    stringsAsFactors = FALSE
  )
  
  # Fill in coefficients from each model
  for (model_name in names(model_list)) {
    model         <- model_list[[model_name]]
    model_summary <- summary(model)
    coefs         <- coef(model_summary)
    
    all_coefs[[model_name]] <- NA
    
    for (i in 1:nrow(all_coefs)) {
      var_name <- all_coefs$Variable[i]
      if (var_name %in% rownames(coefs)) {
        estimate <- coefs[var_name, "Estimate"]
        p_value  <- coefs[var_name, "Pr(>|t|)"]
        all_coefs[i, model_name] <- format_coef(estimate, p_value)
      }
    }
  }
  
  # Build flextable
  n_models <- length(model_list)
  
  ft <- flextable(all_coefs) %>%
    add_header_row(
      values    = c("", "", title_label),
      colwidths = c(1, 1, n_models)
    ) %>%
    theme_booktabs() %>%
    align(j = 1:2,              align = "left") %>%
    align(j = 3:(2 + n_models), align = "center") %>%
    width(j = 1,              width = 3.5) %>%
    width(j = 2,              width = 2.0) %>%
    width(j = 3:(2 + n_models), width = 1.5) %>%
    bold(i = 1, bold = TRUE) %>%
    add_footer_lines(values = c(
      "* p < 0.05, ** p < 0.01, *** p < 0.001",
      "Values show regression coefficients with significance stars",
      "Outcome: Depression symptoms (CBCL DSM-5 T-Score) at 3-year follow-up"
    ))
  
  return(ft)
}

# ── Generate tables ───────────────────────────────────────────────────────────

WH_depress_sample  <- build_coef_table(
  WH_model_depress_full_sample,
  "Depression (CBCL DSM-5 Depression T-Score) — Pre-Matched Sample"
)

WH_depress_matched <- build_coef_table(
  WH_model_depress_full_matched,
  "Depression (CBCL DSM-5 Depression T-Score) — Matched Sample"
)

WH_depress_sample
WH_depress_matched
```


### C3: Substance Use -- Matched (WB)

```{r build-substance-regression-table-wb}
# ── Define model list ─────────────────────────────────────────────────────────
WB_model_subs_full_sample <- list(
  "All Participants" = prematch_SU_model_WB,
  "White Only"       = prematch_SU_model_onlyWB,
  "Black Only"       = prematch_SU_model_onlyB,
  "Race-Controlled"  = prematch_SU_model_race_control_WB
)

WB_model_subs_full_matched <- list(
  "All Participants" = matched_SU_model_WB,
  "White Only"       = matched_SU_model_onlyWB,
  "Black Only"       = matched_SU_model_onlyB,
  "Race-Controlled"  = matched_SU_model_race_control_WB
)

# ── Shared reference levels ───────────────────────────────────────────────────

reference_levels <- c(
  "(Intercept)"                     = "N/A",
  "is_black"                        = "Non-Black (0)",
  "ed_det_susp_baseline_31"         = "No Suspension (0)",
  "tlfb_cal_scr_tot_su_ud_ses-04A"  = "Continuous Score",
  "demo_sex_v22"                    = "Male (1)",
  "crpbi_caregiver1_y1"             = "Continuous Score",
  "income_bracket2"                 = "<$50K (Reference)",
  "income_bracket3"                 = "<$50K (Reference)",
  "income_bracket4"                 = "<$50K (Reference)",
  "income_bracket5"                 = "<$50K (Reference)",
  "income_bracket777"               = "<$50K (Reference)",
  "income_bracket999"               = "<$50K (Reference)",
  "parental_ed_bracket2"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket3"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket4"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket5"            = "Up to HS no diploma (Reference)",
  "marital_status_bracket2"         = "Married/Living Together (Reference)",
  "marital_status_bracket3"         = "Married/Living Together (Reference)",
  "siteFlorida International University"        = "Children's Hospital Los Angeles",
  "siteLaureate Institute for Brain Research"   = "Children's Hospital Los Angeles",
  "siteMedical University of South Carolina"    = "Children's Hospital Los Angeles",
  "siteOregon Health & Science University"      = "Children's Hospital Los Angeles",
  "siteSRI International"                       = "Children's Hospital Los Angeles",
  "siteUniversity of California, Los Angeles"   = "Children's Hospital Los Angeles",
  "siteUniversity of California, San Diego"     = "Children's Hospital Los Angeles",
  "siteUniversity of Colorado Boulder"          = "Children's Hospital Los Angeles",
  "siteUniversity of Florida"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Maryland, Baltimore"       = "Children's Hospital Los Angeles",
  "siteUniversity of Michigan"                  = "Children's Hospital Los Angeles",
  "siteUniversity of Minnesota"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Pittsburgh Medical Center" = "Children's Hospital Los Angeles",
  "siteUniversity of Rochester"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Utah"                      = "Children's Hospital Los Angeles",
  "siteUniversity of Vermont"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Wisconsin, Milwaukee"      = "Children's Hospital Los Angeles",
  "siteVirginia Commonwealth University"        = "Children's Hospital Los Angeles",
  "siteWashington University in St. Louis"      = "Children's Hospital Los Angeles",
  "siteYale University"                         = "Children's Hospital Los Angeles"
)

# ── Core builder function ─────────────────────────────────────────────────────

build_coef_table <- function(model_list, title_label) {
  
  all_variable_names <- get_all_variables(model_list)
  
  # Auto-fill any missing reference levels
  ref_levels <- reference_levels  # copy shared reference levels
  for (var in all_variable_names) {
    if (!(var %in% names(ref_levels))) {
      if      (grepl("income_bracket",         var)) ref_levels[var] <- "<$50K (Reference)"
      else if (grepl("parental_ed_bracket",  var)) ref_levels[var] <- "Up to HS no diploma (Reference)"
      else if (grepl("marital_status_bracket", var)) ref_levels[var] <- "Married/Living Together (Reference)"
      else                                            ref_levels[var] <- "Reference Category"
    }
  }
  
  # Base data frame
  all_coefs <- data.frame(
    Variable  = all_variable_names,
    Reference = ref_levels[all_variable_names],
    stringsAsFactors = FALSE
  )
  
  # Fill in coefficients from each model
  for (model_name in names(model_list)) {
    model         <- model_list[[model_name]]
    model_summary <- summary(model)
    coefs         <- coef(model_summary)
    
    all_coefs[[model_name]] <- NA
    
    for (i in 1:nrow(all_coefs)) {
      var_name <- all_coefs$Variable[i]
      if (var_name %in% rownames(coefs)) {
        estimate <- coefs[var_name, "Estimate"]
        p_value  <- coefs[var_name, "Pr(>|t|)"]
        all_coefs[i, model_name] <- format_coef(estimate, p_value)
      }
    }
  }
  
  # Build flextable
  n_models <- length(model_list)
  
  ft <- flextable(all_coefs) %>%
    add_header_row(
      values    = c("", "", title_label),
      colwidths = c(1, 1, n_models)
    ) %>%
    theme_booktabs() %>%
    align(j = 1:2,              align = "left") %>%
    align(j = 3:(2 + n_models), align = "center") %>%
    width(j = 1,              width = 3.5) %>%
    width(j = 2,              width = 2.0) %>%
    width(j = 3:(2 + n_models), width = 1.5) %>%
    bold(i = 1, bold = TRUE) %>%
    add_footer_lines(values = c(
      "* p < 0.05, ** p < 0.01, *** p < 0.001",
      "Values show regression coefficients with significance stars",
      "Outcome: Substance use at 3-year follow-up"
    ))
  
  return(ft)
}

# ── Generate tables ───────────────────────────────────────────────────────────

WB_subs_sample  <- build_coef_table(
  WB_model_subs_full_sample,
  "Substance Use — Pre-Matched Sample"
)

WB_subs_matched <- build_coef_table(
  WB_model_subs_full_matched,
  "Substance Use — Matched Sample"
)

WB_subs_sample
WB_subs_matched
```

### C4: Substance Use -- Matched (WH)

```{r build-substance-regression-table-wh}
# ── Define model list ─────────────────────────────────────────────────────────
WH_model_subs_full_sample <- list(
  "All Participants" = prematch_SU_model_WH,
  "White Only"       = prematch_SU_model_onlyWH,
  "Hispanic Only"    = prematch_SU_model_onlyH,
  "Race-Controlled"  = prematch_SU_model_race_control_WH
)

WH_model_subs_full_matched <- list(
  "All Participants" = matched_SU_model_WH,
  "White Only"       = matched_SU_model_onlyWH,
  "Hispanic Only"    = matched_SU_model_onlyH,
  "Race-Controlled"  = matched_SU_model_race_control_WH
)

# ── Shared reference levels ───────────────────────────────────────────────────

reference_levels <- c(
  "(Intercept)"                     = "N/A",
  "is_his"                          = "Non-Hispanic (0)",
  "ed_det_susp_baseline_31"         = "No Suspension (0)",
  "tlfb_cal_scr_tot_su_ud_ses-04A"  = "Continuous Score",
  "demo_sex_v22"                    = "Male (1)",
  "crpbi_caregiver1_y1"             = "Continuous Score",
  "income_bracket2"                 = "<$50K (Reference)",
  "income_bracket3"                 = "<$50K (Reference)",
  "income_bracket4"                 = "<$50K (Reference)",
  "income_bracket5"                 = "<$50K (Reference)",
  "income_bracket777"               = "<$50K (Reference)",
  "income_bracket999"               = "<$50K (Reference)",
  "parental_ed_bracket2"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket3"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket4"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket5"            = "Up to HS no diploma (Reference)",
  "marital_status_bracket2"         = "Married/Living Together (Reference)",
  "marital_status_bracket3"         = "Married/Living Together (Reference)",
  "siteFlorida International University"        = "Children's Hospital Los Angeles",
  "siteLaureate Institute for Brain Research"   = "Children's Hospital Los Angeles",
  "siteMedical University of South Carolina"    = "Children's Hospital Los Angeles",
  "siteOregon Health & Science University"      = "Children's Hospital Los Angeles",
  "siteSRI International"                       = "Children's Hospital Los Angeles",
  "siteUniversity of California, Los Angeles"   = "Children's Hospital Los Angeles",
  "siteUniversity of California, San Diego"     = "Children's Hospital Los Angeles",
  "siteUniversity of Colorado Boulder"          = "Children's Hospital Los Angeles",
  "siteUniversity of Florida"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Maryland, Baltimore"       = "Children's Hospital Los Angeles",
  "siteUniversity of Michigan"                  = "Children's Hospital Los Angeles",
  "siteUniversity of Minnesota"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Pittsburgh Medical Center" = "Children's Hospital Los Angeles",
  "siteUniversity of Rochester"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Utah"                      = "Children's Hospital Los Angeles",
  "siteUniversity of Vermont"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Wisconsin, Milwaukee"      = "Children's Hospital Los Angeles",
  "siteVirginia Commonwealth University"        = "Children's Hospital Los Angeles",
  "siteWashington University in St. Louis"      = "Children's Hospital Los Angeles",
  "siteYale University"                         = "Children's Hospital Los Angeles"
)

# ── Core builder function ─────────────────────────────────────────────────────

build_coef_table <- function(model_list, title_label) {
  
  all_variable_names <- get_all_variables(model_list)
  
  # Auto-fill any missing reference levels
  ref_levels <- reference_levels  # copy shared reference levels
  for (var in all_variable_names) {
    if (!(var %in% names(ref_levels))) {
      if      (grepl("income_bracket",         var)) ref_levels[var] <- "<$50K (Reference)"
      else if (grepl("parental_ed_bracket",  var)) ref_levels[var] <- "Up to HS no diploma (Reference)"
      else if (grepl("marital_status_bracket", var)) ref_levels[var] <- "Married/Living Together (Reference)"
      else                                            ref_levels[var] <- "Reference Category"
    }
  }
  
  # Base data frame
  all_coefs <- data.frame(
    Variable  = all_variable_names,
    Reference = ref_levels[all_variable_names],
    stringsAsFactors = FALSE
  )
  
  # Fill in coefficients from each model
  for (model_name in names(model_list)) {
    model         <- model_list[[model_name]]
    model_summary <- summary(model)
    coefs         <- coef(model_summary)
    
    all_coefs[[model_name]] <- NA
    
    for (i in 1:nrow(all_coefs)) {
      var_name <- all_coefs$Variable[i]
      if (var_name %in% rownames(coefs)) {
        estimate <- coefs[var_name, "Estimate"]
        p_value  <- coefs[var_name, "Pr(>|t|)"]
        all_coefs[i, model_name] <- format_coef(estimate, p_value)
      }
    }
  }
  
  # Build flextable
  n_models <- length(model_list)
  
  ft <- flextable(all_coefs) %>%
    add_header_row(
      values    = c("", "", title_label),
      colwidths = c(1, 1, n_models)
    ) %>%
    theme_booktabs() %>%
    align(j = 1:2,              align = "left") %>%
    align(j = 3:(2 + n_models), align = "center") %>%
    width(j = 1,              width = 3.5) %>%
    width(j = 2,              width = 2.0) %>%
    width(j = 3:(2 + n_models), width = 1.5) %>%
    bold(i = 1, bold = TRUE) %>%
    add_footer_lines(values = c(
      "* p < 0.05, ** p < 0.01, *** p < 0.001",
      "Values show regression coefficients with significance stars",
      "Outcome: Substance use at 3-year follow-up"
    ))
  
  return(ft)
}

# ── Generate tables ───────────────────────────────────────────────────────────

WH_subs_sample  <- build_coef_table(
  WH_model_subs_full_sample,
  "Substance Use — Pre-Matched Sample"
)

WH_subs_matched <- build_coef_table(
  WH_model_subs_full_matched,
  "Substance Use — Matched Sample"
)

WH_subs_sample
WH_subs_matched
```


### C4: Academic Grades -- Matched (WB)

```{r build-grades-regression-table-wb}
# ── Define model list ─────────────────────────────────────────────────────────
WB_model_grade_full_sample <- list(
  "All Participants" = prematch_grade_model_WB,
  "White Only"       = prematch_grade_model_onlyWB,
  "Black Only"       = prematch_grade_model_onlyB,
  "Race-Controlled"  = prematch_grade_model_race_control_WB
)

WB_model_grade_full_matched <- list(
  "All Participants" = matched_grade_model_WB,
  "White Only"       = matched_grade_model_onlyWB,
  "Black Only"       = matched_grade_model_onlyB,
  "Race-Controlled"  = matched_grade_model_race_control_WB
)

# ── Shared reference levels ───────────────────────────────────────────────────

reference_levels <- c(
  "(Intercept)"                     = "N/A",
  "is_black"                        = "Non-Black (0)",
  "ed_det_susp_baseline_31"         = "No Suspension (0)",
  "grade_gpa_ses-03A"               = "Continuous Score",
  "grade_gpa_ses-04A"               = "Continuous Score",
  "demo_sex_v22"                    = "Male (1)",
  "crpbi_caregiver1_y1"             = "Continuous Score",
  "income_bracket2"                 = "<$50K (Reference)",
  "income_bracket3"                 = "<$50K (Reference)",
  "income_bracket4"                 = "<$50K (Reference)",
  "income_bracket5"                 = "<$50K (Reference)",
  "income_bracket777"               = "<$50K (Reference)",
  "income_bracket999"               = "<$50K (Reference)",
  "parental_ed_bracket2"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket3"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket4"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket5"            = "Up to HS no diploma (Reference)",
  "marital_status_bracket2"         = "Married/Living Together (Reference)",
  "marital_status_bracket3"         = "Married/Living Together (Reference)",
  "siteFlorida International University"        = "Children's Hospital Los Angeles",
  "siteLaureate Institute for Brain Research"   = "Children's Hospital Los Angeles",
  "siteMedical University of South Carolina"    = "Children's Hospital Los Angeles",
  "siteOregon Health & Science University"      = "Children's Hospital Los Angeles",
  "siteSRI International"                       = "Children's Hospital Los Angeles",
  "siteUniversity of California, Los Angeles"   = "Children's Hospital Los Angeles",
  "siteUniversity of California, San Diego"     = "Children's Hospital Los Angeles",
  "siteUniversity of Colorado Boulder"          = "Children's Hospital Los Angeles",
  "siteUniversity of Florida"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Maryland, Baltimore"       = "Children's Hospital Los Angeles",
  "siteUniversity of Michigan"                  = "Children's Hospital Los Angeles",
  "siteUniversity of Minnesota"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Pittsburgh Medical Center" = "Children's Hospital Los Angeles",
  "siteUniversity of Rochester"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Utah"                      = "Children's Hospital Los Angeles",
  "siteUniversity of Vermont"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Wisconsin, Milwaukee"      = "Children's Hospital Los Angeles",
  "siteVirginia Commonwealth University"        = "Children's Hospital Los Angeles",
  "siteWashington University in St. Louis"      = "Children's Hospital Los Angeles",
  "siteYale University"                         = "Children's Hospital Los Angeles"
)

# ── Core builder function ─────────────────────────────────────────────────────

build_coef_table <- function(model_list, title_label) {
  
  all_variable_names <- get_all_variables(model_list)
  
  # Auto-fill any missing reference levels
  ref_levels <- reference_levels  # copy shared reference levels
  for (var in all_variable_names) {
    if (!(var %in% names(ref_levels))) {
      if      (grepl("income_bracket",         var)) ref_levels[var] <- "<$50K (Reference)"
      else if (grepl("parental_ed_bracket",  var)) ref_levels[var] <- "Up to HS no diploma (Reference)"
      else if (grepl("marital_status_bracket", var)) ref_levels[var] <- "Married/Living Together (Reference)"
      else                                            ref_levels[var] <- "Reference Category"
    }
  }
  
  # Base data frame
  all_coefs <- data.frame(
    Variable  = all_variable_names,
    Reference = ref_levels[all_variable_names],
    stringsAsFactors = FALSE
  )
  
  # Fill in coefficients from each model
  for (model_name in names(model_list)) {
    model         <- model_list[[model_name]]
    model_summary <- summary(model)
    coefs         <- coef(model_summary)
    
    all_coefs[[model_name]] <- NA
    
    for (i in 1:nrow(all_coefs)) {
      var_name <- all_coefs$Variable[i]
      if (var_name %in% rownames(coefs)) {
        estimate <- coefs[var_name, "Estimate"]
        p_value  <- coefs[var_name, "Pr(>|t|)"]
        all_coefs[i, model_name] <- format_coef(estimate, p_value)
      }
    }
  }
  
  # Build flextable
  n_models <- length(model_list)
  
  ft <- flextable(all_coefs) %>%
    add_header_row(
      values    = c("", "", title_label),
      colwidths = c(1, 1, n_models)
    ) %>%
    theme_booktabs() %>%
    align(j = 1:2,              align = "left") %>%
    align(j = 3:(2 + n_models), align = "center") %>%
    width(j = 1,              width = 3.5) %>%
    width(j = 2,              width = 2.0) %>%
    width(j = 3:(2 + n_models), width = 1.5) %>%
    bold(i = 1, bold = TRUE) %>%
    add_footer_lines(values = c(
      "* p < 0.05, ** p < 0.01, *** p < 0.001",
      "Values show regression coefficients with significance stars",
      "Outcome: GPA at 3-year follow-up"
    ))
  
  return(ft)
}

# ── Generate tables ───────────────────────────────────────────────────────────

WB_grade_sample  <- build_coef_table(
  WB_model_grade_full_sample,
  "Academic Grade — Pre-Matched Sample"
)

WB_grade_matched <- build_coef_table(
  WB_model_grade_full_matched,
  "Academic Grade — Matched Sample"
)

WB_grade_sample
WB_grade_matched
```


### C5: Academic Grade -- Matched (WH)

```{r build-grades-regression-table-wh}
# ── Define model list ─────────────────────────────────────────────────────────
WH_model_grade_full_sample <- list(
  "All Participants" = prematch_grade_model_WH,
  "White Only"       = prematch_grade_model_onlyWH,
  "Hispanic Only"    = prematch_grade_model_onlyH,
  "Race-Controlled"  = prematch_grade_model_race_control_WH
)

WH_model_grade_full_matched <- list(
  "All Participants" = matched_grade_model_WH,
  "White Only"       = matched_grade_model_onlyWH,
  "Hispanic Only"    = matched_grade_model_onlyH,
  "Race-Controlled"  = matched_grade_model_race_control_WH
)

# ── Shared reference levels ───────────────────────────────────────────────────

reference_levels <- c(
  "(Intercept)"                     = "N/A",
  "is_his"                          = "Non-Hispanic (0)",
  "ed_det_susp_baseline_31"         = "No Suspension (0)",
  "grade_gpa_ses-03A"               = "Continuous Score",
  "grade_gpa_ses-04A"               = "Continuous Score",
  "demo_sex_v22"                    = "Male (1)",
  "crpbi_caregiver1_y1"             = "Continuous Score",
  "income_bracket2"                 = "<$50K (Reference)",
  "income_bracket3"                 = "<$50K (Reference)",
  "income_bracket4"                 = "<$50K (Reference)",
  "income_bracket5"                 = "<$50K (Reference)",
  "income_bracket777"               = "<$50K (Reference)",
  "income_bracket999"               = "<$50K (Reference)",
  "parental_ed_bracket2"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket3"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket4"            = "Up to HS no diploma (Reference)",
  "parental_ed_bracket5"            = "Up to HS no diploma (Reference)",
  "marital_status_bracket2"         = "Married/Living Together (Reference)",
  "marital_status_bracket3"         = "Married/Living Together (Reference)",
  "siteFlorida International University"        = "Children's Hospital Los Angeles",
  "siteLaureate Institute for Brain Research"   = "Children's Hospital Los Angeles",
  "siteMedical University of South Carolina"    = "Children's Hospital Los Angeles",
  "siteOregon Health & Science University"      = "Children's Hospital Los Angeles",
  "siteSRI International"                       = "Children's Hospital Los Angeles",
  "siteUniversity of California, Los Angeles"   = "Children's Hospital Los Angeles",
  "siteUniversity of California, San Diego"     = "Children's Hospital Los Angeles",
  "siteUniversity of Colorado Boulder"          = "Children's Hospital Los Angeles",
  "siteUniversity of Florida"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Maryland, Baltimore"       = "Children's Hospital Los Angeles",
  "siteUniversity of Michigan"                  = "Children's Hospital Los Angeles",
  "siteUniversity of Minnesota"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Pittsburgh Medical Center" = "Children's Hospital Los Angeles",
  "siteUniversity of Rochester"                 = "Children's Hospital Los Angeles",
  "siteUniversity of Utah"                      = "Children's Hospital Los Angeles",
  "siteUniversity of Vermont"                   = "Children's Hospital Los Angeles",
  "siteUniversity of Wisconsin, Milwaukee"      = "Children's Hospital Los Angeles",
  "siteVirginia Commonwealth University"        = "Children's Hospital Los Angeles",
  "siteWashington University in St. Louis"      = "Children's Hospital Los Angeles",
  "siteYale University"                         = "Children's Hospital Los Angeles"
)

# ── Core builder function ─────────────────────────────────────────────────────

build_coef_table <- function(model_list, title_label) {
  
  all_variable_names <- get_all_variables(model_list)
  
  # Auto-fill any missing reference levels
  ref_levels <- reference_levels  # copy shared reference levels
  for (var in all_variable_names) {
    if (!(var %in% names(ref_levels))) {
      if      (grepl("income_bracket",         var)) ref_levels[var] <- "<$50K (Reference)"
      else if (grepl("parental_ed_bracket",  var)) ref_levels[var] <- "Up to HS no diploma (Reference)"
      else if (grepl("marital_status_bracket", var)) ref_levels[var] <- "Married/Living Together (Reference)"
      else                                            ref_levels[var] <- "Reference Category"
    }
  }
  
  # Base data frame
  all_coefs <- data.frame(
    Variable  = all_variable_names,
    Reference = ref_levels[all_variable_names],
    stringsAsFactors = FALSE
  )
  
  # Fill in coefficients from each model
  for (model_name in names(model_list)) {
    model         <- model_list[[model_name]]
    model_summary <- summary(model)
    coefs         <- coef(model_summary)
    
    all_coefs[[model_name]] <- NA
    
    for (i in 1:nrow(all_coefs)) {
      var_name <- all_coefs$Variable[i]
      if (var_name %in% rownames(coefs)) {
        estimate <- coefs[var_name, "Estimate"]
        p_value  <- coefs[var_name, "Pr(>|t|)"]
        all_coefs[i, model_name] <- format_coef(estimate, p_value)
      }
    }
  }
  
  # Build flextable
  n_models <- length(model_list)
  
  ft <- flextable(all_coefs) %>%
    add_header_row(
      values    = c("", "", title_label),
      colwidths = c(1, 1, n_models)
    ) %>%
    theme_booktabs() %>%
    align(j = 1:2,              align = "left") %>%
    align(j = 3:(2 + n_models), align = "center") %>%
    width(j = 1,              width = 3.5) %>%
    width(j = 2,              width = 2.0) %>%
    width(j = 3:(2 + n_models), width = 1.5) %>%
    bold(i = 1, bold = TRUE) %>%
    add_footer_lines(values = c(
      "* p < 0.05, ** p < 0.01, *** p < 0.001",
      "Values show regression coefficients with significance stars",
      "Outcome: GPA at 3-year follow-up"
    ))
  
  return(ft)
}

# ── Generate tables ───────────────────────────────────────────────────────────

WH_grade_sample  <- build_coef_table(
  WH_model_grade_full_sample,
  "Academic Grade — Pre-Matched Sample"
)

WH_grade_matched <- build_coef_table(
  WH_model_grade_full_matched,
  "Academic Grade — Matched Sample"
)

WH_grade_sample
WH_grade_matched
```


```{r save-regression-table}
# ============================================================
# Save the regression table to a Word document.
# ============================================================

# Depression 
# WB
save_as_docx(
  `WB_Depression_PreMatch_Regression` = WB_depress_sample,
  path = "output/WB_Depression_PreMatch_Regression.docx"
)
cat("Saved: WB_Depression_PreMatch_Regression.docx\n")

save_as_docx(
  `WB_Depression_Matched_Regression` = WB_depress_matched,
  path = "output/WB_Depression_Matched_Regression.docx"
)
cat("Saved: WB_Depression_Matched_Regression.docx\n")

# WH
save_as_docx(
  `WH_Depression_PreMatch_Regression` = WH_depress_sample,
  path = "output/WH_Depression_PreMatch_Regression.docx"
)
cat("Saved: WH_Depression_PreMatch_Regression.docx\n")

save_as_docx(
  `WH_Depression_Matched_Regression` = WH_depress_matched,
  path = "output/WH_Depression_Matched_Regression.docx"
)
cat("Saved: WH_Depression_Matched_Regression.docx\n")

# Academic grades
# WB
save_as_docx(
  `WB_GPA_PreMatch_Regression` = WB_grade_sample,
  path = "output/WB_GPA_PreMatch_Regression.docx"
)
cat("Saved: WB_GPA_PreMatch_Regression.docx\n")

save_as_docx(
  `WB_GPA_Matched_Regression` = WB_grade_matched,
  path = "output/WB_GPA_Matched_Regression.docx"
)
cat("Saved: WB_GPA_Matched_Regression.docx\n")

# WH
save_as_docx(
  `WH_GPA_PreMatch_Regression` = WH_grade_sample,
  path = "output/WH_GPA_PreMatch_Regression.docx"
)
cat("Saved: WH_GPA_PreMatch_Regression.docx\n")

save_as_docx(
  `WH_GPA_Matched_Regression` = WH_grade_matched,
  path = "output/WH_GPA_Matched_Regression.docx"
)
cat("Saved: WB_GPA_Matched_Regression.docx\n")
```

# Analysis of Variance

Examine the relationship between rule-breaking behavior (`cbcl_scr_syn_rulebreak_t`) and the proportion of students receiving disciplinary action (`ed_det_susp_baseline_3`), and whether this relationship varied by race/ethnicity.

::: callout-note
ANOVA requires a continuous dependent variable (rulebreaking t-score) and an independent variable with at least 2 levels which determine the groups to compare.

Normally you should test for normality visually using a histogram and a Q-Q plot; however, in case of large samples, [normality is not required]{.underline} due to the central limit theorem. You should test for the equality of variances using a boxplot for example or formally via the Levene's test or Bartlett's test. There should be no significant outliers.
:::

## Equality of Variances

```{r anova-wb-matched}
EV_WB <- WB_matched_data_raw %>%
  mutate(is_black = factor(is_black, labels = c("White", "Black"))) %>%
  ggplot( aes(x=is_black, y=cbcl_scr_syn_rulebreak_t, fill=is_black)) +
    geom_violin(alpha = 0.6) +
    scale_fill_brand(picks = c(1, 20)) +
    theme_brand() + labs(y = "Rule-breaking t-score") +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("A Boxplot of Rulebreak T-Score by Race/Ethnicity (White vs Black)") +
    xlab("")
EV_WB

```

```{r anova-wh-matched}
EV_WH <- WH_matched_data_raw %>%
  mutate(is_his = factor(is_his, labels = c("White", "Hispanic"))) %>%
  ggplot( aes(x=is_his, y=cbcl_scr_syn_rulebreak_t, fill=is_his)) +
    geom_violin(alpha = 0.6) +
    scale_fill_brand(picks = c(1, 20)) +
    theme_brand() + labs(y = "Rule-breaking t-score") +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("A Boxplot of Rulebreak T-Score by Race/Ethnicity (White vs Hispanic)") +
    xlab("")
EV_WH

equality_var <- cowplot::plot_grid(EV_WB, EV_WH, nrow = 2)

ggsave("output/equality_var.png",
       equality_var, width = 6, height = 6)
```

The dispersion is more or less the same between the racial groups.

## Rule Breaking Descriptives

Descriptives for rule breaking behavior at baseline by race/ethnicity.

```{r descript-rulebreaking}
# relabel variables for readability
filtered_df_baseline_recoded_W_anov <- filtered_df_baseline_recoded_W %>%
  mutate(`CBCL Rule-Break T-score` = cbcl_scr_syn_rulebreak_t, 
         `Breaks Rules (CBCL q28)` = factor(cbcl_q28_p, labels = c("Not True", "Somewhat/Sometimes True", "Very True/Often True"))
  )

filtered_df_baseline_recoded_B_anov <- filtered_df_baseline_recoded_B %>%
  mutate(`CBCL Rule-Break T-score` = cbcl_scr_syn_rulebreak_t, 
         `Breaks Rules (CBCL q28)` = factor(cbcl_q28_p, labels = c("Not True", "Somewhat/Sometimes True", "Very True/Often True"))
  )

filtered_df_baseline_recoded_H_anov <- filtered_df_baseline_recoded_H %>%
  mutate(`CBCL Rule-Break T-score` = cbcl_scr_syn_rulebreak_t, 
         `Breaks Rules (CBCL q28)` = factor(cbcl_q28_p, labels = c("Not True", "Somewhat/Sometimes True", "Very True/Often True"))
  )

# create tables 
table_rule_white <- table1(
  ~ `CBCL Rule-Break T-score` + `Breaks Rules (CBCL q28)`,
  data = filtered_df_baseline_recoded_W_anov,
  caption = "Rule-Breaking by Race/Ethnicity at Baseline: White Participants"
  #render.missing = NULL
)
table_rule_white

table_rule_black <- table1(
  ~ `CBCL Rule-Break T-score` + `Breaks Rules (CBCL q28)`,
  data = filtered_df_baseline_recoded_B_anov,
  caption = "Rule-Breaking by Race/Ethnicity at Baseline: Black Participants"
  #render.missing = NULL
)
table_rule_black

table_rule_his <- table1(
  ~ `CBCL Rule-Break T-score` + `Breaks Rules (CBCL q28)`,
  data = filtered_df_baseline_recoded_H_anov,
  caption = "Rule-Breaking by Race/Ethnicity at Baseline: Hispanic Participants"
  #render.missing = NULL
)
table_rule_his
```

## D1: One-Way ANOVA for T-Scores (Continous)

OUTCOME: suspension by year 3 

PREDICTORS: CBCL rule breaking (t-score) and race/ethnicity

```{r anova-black}
res_aov_B <- aov(ed_det_susp_baseline_3 ~ cbcl_scr_syn_rulebreak_t * as.factor(is_black), 
  data = WB_matched_data_raw,
) # <3>

#summary(res_aov_B)
report(res_aov_B)
```

3.  The asterisk gives you both main effects and the interaction in one call, equivalent to writing a + b + a:b

```{r anova-his}
res_aov_H <- aov(ed_det_susp_baseline_3 ~ cbcl_scr_syn_rulebreak_t * as.factor(is_his),
  data = WH_matched_data_raw,
)

#summary(res_aov_H)
report(res_aov_H)
```

## D2: One-Way ANOVA for T-Scores (Quartiles)

OUTCOME: suspension by year 3 

PREDICTORS: CBCL rule breaking (t-score in quartiles) and race/ethnicity

```{r anova-black-quart}
WB_matched_data_raw <- WB_matched_data_raw %>%
  mutate(cbcl_scr_syn_rulebreak_t_quart = ntile(cbcl_scr_syn_rulebreak_t, 4))

res_aov_B_quart <- aov(ed_det_susp_baseline_3 ~ cbcl_scr_syn_rulebreak_t_quart * as.factor(is_black), 
  data = WB_matched_data_raw,
)

#summary(res_aov_B_quart)
report(res_aov_B_quart)
```

```{r anova-his-quart}
WH_matched_data_raw <- WH_matched_data_raw %>%
  mutate(cbcl_scr_syn_rulebreak_t_quart = ntile(cbcl_scr_syn_rulebreak_t, 4))

res_aov_H_quart <- aov(ed_det_susp_baseline_3 ~ cbcl_scr_syn_rulebreak_t_quart * as.factor(is_his),
  data = WH_matched_data_raw,
)

#summary(res_aov_H_quart)
report(res_aov_H_quart)
```

# Build Visualizations

## Rulebreaking Quartiles

White-Black matched sample
```{r rulebreak-bar-wb, warning=FALSE}
abcd_all <- abcd_all %>%
  mutate(cbcl_scr_syn_rulebreak_t_quart = ntile(cbcl_scr_syn_rulebreak_t, 4)) %>%
  mutate(demo_sex_v2 = factor(demo_sex_v2),
         marital_status_bracket = factor(marital_status_bracket),
         time = factor(eventname, labels = c("Baseline", "1 Year", "2 Year", "3 Year", "4 Year", "5 Year", "6 Year", "7 Year")))

WB_matched_abcd_all <- semi_join(abcd_all, WB_matched_data, by = "src_subject_id")

WB_matched_abcd_all <- WB_matched_abcd_all %>%
  mutate(
    race_ethnicity_recode = case_when(
      race_ethnicity == 2         ~ "White",
      race_ethnicity %in% c(3, 6) ~ "Black",
      TRUE                        ~ NA_character_
    ),
    race_ethnicity_recode = factor(race_ethnicity_recode, levels = c("White", "Black"))
  ) %>%
  filter(!is.na(race_ethnicity_recode))

n_distinct(WB_matched_data_raw$src_subject_id)
n_distinct(WB_matched_abcd_all$src_subject_id)
cat("White participants (control):",
    n_distinct(WB_matched_abcd_all$src_subject_id[WB_matched_abcd_all$race_ethnicity_recode == "White"]),
    "\n")

cat("Black participants (treated):",
    n_distinct(WB_matched_abcd_all$src_subject_id[WB_matched_abcd_all$race_ethnicity_recode == "Black"]),
    "\n")

WB_rulebreak_quart_plot <- WB_matched_abcd_all %>%
  filter(time == "Baseline" | time == "1 Year" | time == "2 Year" | time == "3 Year" | time == "4 Year") %>%
  filter(!is.na(cbcl_scr_syn_rulebreak_t_quart)) %>%
  ggplot(aes(x = factor(cbcl_scr_syn_rulebreak_t_quart), y = kbi_p_c_det_susp_l, fill = race_ethnicity_recode)) +
  stat_summary(
    fun = "mean",
    geom = "col",
    position = position_dodge(width = 0.9),
    color = "black", width = 0.8
  ) +
  stat_summary(
    fun.data = "mean_se",   # or mean_cl_normal for 95% CI
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  facet_wrap(~ time) +
  scale_fill_manual(values = c("White" = "#D16284", "Black" = "#407088")) +
  labs(x = "Rule-breaking t-score (quartiles)", y = "Proportion receiving disciplinary action", fill = "Ethnoracial identity", title = "Proportion of disciplinary action by rule-breaking t-score\n in the White (n=3421) and Black (n=1340)\n matched sample") +
  theme_brand() + theme(plot.title = element_text(face = "bold", hjust = 0.5))
WB_rulebreak_quart_plot

ggsave("output/WB_rulebreak_quart_plot.png",
       WB_rulebreak_quart_plot, width = 9, height = 8)
```

White-Hispanic matched sample
```{r rulebreak-bar-wh, warning=FALSE}
WH_matched_abcd_all <- semi_join(abcd_all, WH_matched_data, by = "src_subject_id")

WH_matched_abcd_all <- WH_matched_abcd_all %>%
  mutate(
      race_ethnicity_recode = case_when(
      race_ethnicity == 2           ~ "White",
      race_ethnicity %in% c(1, 5)   ~ "Latine",
      TRUE                          ~ NA_character_
    ),
    race_ethnicity_recode = factor(race_ethnicity_recode, levels = c("White", "Latine"))
  ) %>%
  filter(!is.na(race_ethnicity_recode))

n_distinct(WH_matched_data$src_subject_id)
n_distinct(WH_matched_abcd_all$src_subject_id)
cat("White participants (control):",
    n_distinct(WH_matched_abcd_all$src_subject_id[WH_matched_abcd_all$race_ethnicity_recode == "White"]),
    "\n")

cat("Latine participants (treated):",
    n_distinct(WH_matched_abcd_all$src_subject_id[WH_matched_abcd_all$race_ethnicity_recode == "Latine"]),
    "\n")

WH_rulebreak_quart_plot <- WH_matched_abcd_all %>%
  filter(time == "Baseline" | time == "1 Year" | time == "2 Year" | time == "3 Year" | time == "4 Year") %>%
  filter(!is.na(cbcl_scr_syn_rulebreak_t_quart)) %>%
  ggplot(aes(x = factor(cbcl_scr_syn_rulebreak_t_quart), y = kbi_p_c_det_susp_l, fill = race_ethnicity_recode)) +
  stat_summary(
    fun = "mean",
    geom = "col",
    position = position_dodge(width = 0.9),
    color = "black", width = 0.8
  ) +
  stat_summary(
    fun.data = "mean_se",   # or mean_cl_normal for 95% CI
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  facet_wrap(~ time) +
  scale_fill_manual(values = c("White" = "#D16284", "Latine" = "#6E4499")) +
  labs(x = "Rule-breaking t-score (quartiles)", y = "Proportion receiving disciplinary action", fill = "Ethnoracial identity", title = "Proportion of disciplinary action by rule-breaking t-score\n in the White (n=4319) and Latine (n=1733)\n matched sample") +
  theme_brand() + theme(plot.title = element_text(face = "bold", hjust = 0.5))
WH_rulebreak_quart_plot

ggsave("output/WH_rulebreak_quart_plot.png",
       WH_rulebreak_quart_plot, width = 9, height = 8)

ggsave("output/combined_rulebreak_quart_plot.png",
       cowplot::plot_grid(WB_rulebreak_quart_plot, WH_rulebreak_quart_plot, nrow = 2), 
       width = 9, height = 14)
```

# Clean up R workspace

```{r, results='hide'}
gc()
rm(list = ls())
cat("\014")  
```
